### 2024-10-26 docs/paper.qmd
Pre SSRN

---
title: |
  | How does the Accounting Research Community
  | Think About Open Science?^[Joachim Gassen is based at Humboldt-Universität zu Berlin while Jacqueline Klug and Victor van Pelt are affiliated with WHU Otto Beisheim School of Management. We thank Kris Hardies as well as audiences at University of Antwerp, BI Norwegian Business School, Humboldt University Berlin, Stanford University, and WHU for providing helpful feedback. Scott Summers and David Wood thankfully provided data from the BYU Accounting Research Rankings. Excellent research assistance by Benedikt Hahn, Fabian Kalife, and Timnah Weckner is gratefully acknowledged. This work was supported by Deutsche Forschungsgemeinschaft (Project-ID 403041268, TRR 266 Accounting for Transparency).]
author: 
  - name: Joachim Gassen
    email: gassen@wiwi.hu-berlin.de
    affiliation: Humboldt-Universität zu Berlin
  - name: Jacqueline Klug
    email: Jacqueline.Klug@whu.edu
    affiliation: WHU Otto Beisheim School of Management
  - name: Victor van Pelt
    email: Victor.vanPelt@whu.edu 
    affiliation: WHU Otto Beisheim School of Management
date: last-modified
date-format: long
execute:
  echo: false
  message: false
  warning: false
format:
  pdf:
    documentclass: article
    number-sections: true
    colorlinks: true
    papersize: letter
    fig-pos: H
    fig_caption: yes
    fig-cap-location: top
    geometry: margin=1in
    fontsize: 11pt
    ident: yes

editor: visual

bibliography: references.bib
biblio-style: apsr

always_allow_html: yes

header-includes:
  - \usepackage{setspace}\doublespacing
  - \setlength{\parindent}{4em}
  - \setlength{\parskip}{0em}
  - \setlength{\skip\footins}{2em}
  - \usepackage[hang]{footmisc}
  - \setlength{\footnotemargin}{1em}
  - \usepackage{pdfpages}
  - \usepackage{pdflscape}
  - \usepackage{csquotes}
  - \usepackage{lscape}
  - \usepackage{graphicx}
  - \usepackage{pdflscape}
  - \renewcommand{\mkbegdispquote}[2]{\itshape}
  - \usepackage{longtable}
  - \usepackage{geometry}
  - \geometry{a4paper, margin=1in}
  - \usepackage{setspace}
---

```{r setup, include=FALSE}
library(knitr)
library(rmarkdown)
library(gt)
library(modelsummary)
# opts_chunk$set(fig.pos = 'p') # Places figures on their own pages
opts_chunk$set(out.width = '80%', dpi=600)
opts_chunk$set(echo=FALSE, message=FALSE, warning=FALSE, cache = FALSE)

source("../code/table_functions.R")
source("../code/figure_functions.R")

sd <- readRDS("../data/generated/survey_response.rds")
mr <- readRDS("../data/generated/merged_response.rds")
vars <- read_csv("../data/external/variables.csv", show_col_types = FALSE)
pop <- read_csv("../data/external/invited_authors_anon.csv", show_col_types = FALSE)
byu <- read_csv("../data/external/byu_author_data_anon.csv", show_col_types = FALSE)

authors <- read_csv("../data/external/obs_data_authors_anon.csv", show_col_types = FALSE)
papers <- read_csv("../data/external/obs_data_papers_anon.csv", show_col_types = FALSE)

source("../code/obs_data_analyses.R")

gt_add_header_mpage_table <- function(x) {
  x <- sub("\\\\midrule", "\\\\midrule \\\\endhead", x)
  x
}

```

```{=tex}
\thispagestyle{empty}
\begin{abstract}
\singlespace
To understand how the accounting research community thinks about open science, we survey a sample of scholars who have published in leading accounting journals. Building on a voluntary disclosure-themed discussion of the costs and benefits of open accounting research, we document that accounting researchers tend to be more skeptical about the reproducibility of their influential findings than researchers from other social science areas are about theirs. Also, they are less familiar with the public sharing of research materials and pre-register their study designs less often. While they believe that accounting researchers in general and editors in particular favor research material sharing, they are skeptical about its actual prevalence in the accounting research community. We contrast these survey results with observational data, documenting that research material sharing is indeed relatively rare and often linked to journals with data and code sharing policies. We interpret this descriptive evidence as indicating an "accounting open science dilemma" and suggest strategies for addressing it.
\end{abstract}
```
```{=tex}
\vspace{0.5cm}
\begin{flushleft}
\textbf{Keywords}: Accounting Research, Open Science, Reproducibility, Research Material Sharing, Pre-Registration, Survey \newline
\textbf{Data Availability}: Data available in code repository \newline
\textbf{Code Repository}: \href{https://github.com/joachim-gassen/osacc}{https://github.com/joachim-gassen/osacc}  \newline
\textbf{Survey Dashboard}: \href{https://jgassen.shinyapps.io/osacc/}{https://jgassen.shinyapps.io/osacc} \newline
\textbf{Declaration of Interest}: The author(s) declare(s) that they have no conflict of interest \newline
\end{flushleft}
```
\pagebreak

\setcounter{page}{1}

## Introduction

Cumulative science builds on the reproducibility of research, which is the ability to recreate the results of prior studies using the same materials as were used by the original researchers [@NSF_2015].[^1] Across all fields of science [@B_2016; @CDF_2016; @M_2017] and within accounting [@HLL_2000; @G_2023; @S_2023; @CGL_2023], scholars have diagnosed shortcomings in reproducibility and advocated the use of various open science methods. Open science implies conducting research in ways that allow others to collaborate and contribute. Two central open science methods that have been put forward as strategies to improve the reproducibility of our work are the public sharing of research materials and the pre-registration of study plans. Over the last years, there has been a significant uptake of these open sciences methods in social science areas outside of accounting [@FLCPSWMBP_2023].

Research material sharing and pre-registrations can be understood as disclosure mechanisms that allow researchers to contribute public goods to the research community while also signaling the quality of their work. However, limited individual benefits and significant direct as well as indirect costs make it likely that researchers, on average, will rationally under-engage in these activities, potentially resulting in a socially sub-optimal under-provision of public research materials and on overall lacking research transparency. Following these thoughts and building on prior literature in neighboring fields, we conduct a survey among accounting researchers to explore the perceptions and attitudes towards open science in our field.

Based on the responses of `r nrow(sd)` published accounting researchers, we find that attitudes towards open science clearly differ from those of other social science researchers. Most prominently, accounting researchers tend to hold a generally more pessimistic view of the reproducibility of influential work within their field. Compared to economists and sociologists, for example, accounting researchers are about half a point more skeptical on a 5-point Likert scale. Exploring this assessment across the respondents' demographics, we find this skepticism to be relatively stable. Analytical and qualitative researchers are somewhat more pessimistic. Analytical researchers tend to be particularly more critical of research that uses non-analytical methods.

Given this finding, one would expect accounting researchers to take a proactive stance and embrace two prominent open science methods that are expected to increase the reproducibility of research: The public sharing of research materials (i.e., data, code and research procedures) and the pre-registration of research plans. Turning first to the public sharing of research materials, we document that, again compared to researchers from other social science areas, accounting researchers are somewhat less familiar with this activity and assess it to be somewhat less important. Familiarity is largest for respondents using archival, field, and experimental methods and lowest for respondents using qualitative methods, who also perceive the sharing of research materials as less important.

Similar to, albeit even more pronounced than, other social science researchers, there is a stark contrast in accounting researchers' perceived (prescriptive) importance of research material sharing, their stated willingness to share research materials, and their (descriptive) perception about actual sharing behavior in the accounting research community. 57% of our respondents state that they, at least sometimes, publicly share their research materials. In addition, our respondents believe that about 44% of them favor publicly sharing research materials, and they believe 61% of the top-six accounting editors to be in favor of publicly sharing research materials. Although this sounds comforting and consistent with a general supportive stance of the accounting community toward research material sharing, the percentage comes down to a much more sobering 28% when respondents are asked about actual sharing behavior in the accounting research community. This assessment and the gap it reveals differ significantly across methodological areas. Analytical researchers have the most narrow gap (they expect 43% of accounting researchers to be in favor and 38% to actually share research materials). Archival researchers, in turn, have the widest gap (45% expected to be in favor versus 24% expected to actually share). The overall lowest numbers are again reported by qualitative researchers (36% versus 18%).

Regarding pre-registration, accounting researchers seem as familiar with the concept as other social science researchers. However, our respondents report that they started using it later than other social science researchers (roughly two years) and that they use it less frequently (17% versus 27%). Their trust in pre-registered work is generally comparable with other social science researchers but significantly lower than the trust levels of researchers in Psychology (the field that reports the highest usage rate of pre-registration). A detailed look into differences across methodological fields also reflects this, as we find that experimental researchers use pre-registration most (28%) and express the strongest belief that it increases the trustworthiness of research.

To contrast these perceptions from our survey with actual behavior, we gather additional observational data from a country-stratified random sample of `r nrow(authors)` authors of our target population. We find evidence for at least some form of research material sharing for 39% of the authors in this sample. This number is similar to the self-reported perception of actual research material sharing in our survey (i.e., 38%). However, when we inspect the observational data further, we find that research material sharing in accounting is largely an occasional compliance activity. For instance for 7.5% of the papers written by the authors in the observational data show evidence for research material sharing. Also, when we exclude Journal of Accounting Research and Management Science, two journals with data and code sharing policies, the share of authors for which we document at least some form of research material sharing drops from 39% to 23%. Furthermore, we document very few pre-registrations in our observational data, and all are linked to the Registered Report Initiative of the Journal of Accounting Research.

In light of these numbers, a logical next step is to explore the main rationales for and against sharing research materials. We do so based on a short experimental design using a hypothetical scenario and by using our survey data.[^2] In our hypothetical scenario involving a fellow researcher requesting access to participants' research materials, our participants state that they would be very likely to share on average (mean probability is 72%). We find no evidence that this probability is affected by the randomized nature of the provided rationale (i.e., collaboration or control) and the provided scope of the sharing decision (public versus private sharing). Based on the verbal response of our participants, researchers tend to share their materials as they believe that this improves the quality of their research and of research in general. Also, they share because they want to support their fellow researchers. The most commonly stated reasons for not sharing research materials are the proprietary nature of the data used (56% of respondents give this reason), the time it takes to make materials available (30%), and the fear of being scooped when materials become publicly available (23%). The reasons differ across methods. For qualitative researchers, deidentification problems are very prominent (46%), and analytical researchers consider the time it takes to make materials available far less of a reason than researchers using other methods (i.e., 9.5% versus 30% overall).

Taken together, our findings indicate somewhat of an "accounting open science dilemma": We accounting researchers tend to be skeptical about both the reproducibility of our work, and the ability of research material sharing and pre-registration to improve it. This is consistent with these tools currently being less established in the accounting domain than in other fields. The challenges that hinder reproducibility and the potential effectiveness of open science methods vary across methodological paradigms. Qualitative research faces different challenges than archival quantitative work, and they both differ again from experimental studies. We conclude by discussing these challenges and providing suggestions on how our diagnosed dilemma might be overcome.

## Expectations and Study Design

### Costs and Benefits of Open Accounting Research

Accounting is a field within the social sciences. As such, many of the arguments that have been developed with regard to the costs and benefits of open science in general [@AM_2019; @ABC_2021; @G_2022; @FLCPSWMBP_2023] also apply to the field of accounting. In the following, we present and extend some of these thoughts in a voluntary disclosure framework [@BCLW_2010]. For this purpose, our discussion will be primarily anchored on research material sharing (the main focus of our study) but should largely extend to pre-registration. Understood as a disclosure activity related to underlying research, publicly available research materials constitute a public good. To the extent that they are reusable, they accelerate scientific discoveries, improving the quantity and very likely also the quality of research. Even if they are not really reusable or reusing them would be prohibitively costly, they still reduce the costs of reproducing the original work, thereby reducing assurance costs. As the work in auditing [e.g., @S_1980] posits, reduced auditing costs (easier reproduction) should lead to an overall increase in audit quality (more and better reproductions) resulting in better financial reporting quality (better research).

This means that similar to other disclosure activities, sharing research materials does not necessarily benefit the producer but the research community at large. While intrinsic motivation for good research can be expected to be a strong determinant for individuals self-selecting into the academic accounting profession, one can still expect accounting researchers to under-engage in research material sharing because the production costs are mostly private while the benefits are mostly public. So, are there any private benefits of research material sharing? Besides the warm glow of helping others, research material sharing likely improves the quality of the sharing individual's research output. First, it increases the likelihood that mistakes, conditionally on being present, are being detected. As it seems safe to assume that such incidents impose a cost on sharing researchers, this should increase their incentives to increase research quality prior to sharing. Second, if research materials are shared relatively early, the sharing researcher can directly benefit from community feedback prior to the ultimate publication of the study. Finally, uncovered mistakes are an important learning opportunity, likely improving research quality in the long run.

Based on these arguments, there are reasons to believe that voluntary research material sharing can act as a signal for research quality, creating an incentive for researchers to use it to signal the quality of their work in the publishing and career process. In ideal circumstances, we should obtain the well-known unraveling result where researchers even share and improve low-quality research because the research community would otherwise assume the worst [@M_1981]. One critical obstacle to the unraveling result are disclosure costs [@V_1983], also known as proprietary costs in the contemporary disclosure literature. Thus, for research materials sharing to be viable it is important that the costs of sharing are sufficiently low, especially for researchers that produce high quality work. Thus, it critically hinges on the above quality assurance process. Also, it depends on the direct and indirect costs of research material sharing to be bearable and to be at least comparable for high and low quality researchers.

The direct costs of research material sharing can be sizable. Preparing data, code and research procedures for dissemination requires time. While this time can be expected to also have a quality improvement effect (see above), at least some of the documentation work and code refactoring to increase accessibility and reusability creates costs that will not directly pay off on the quality of the underlying work. Another important factor to consider is the cost of maintaining publicly available materials. While some replication packages or similar materials are static in nature and require only limited maintenance, data and code packages can require ongoing updating and support activities. One important aspect of maintenance costs is that they increase significantly with public demand. As demand is directly linked to the scientific impact of the underlying work it can be argued that high maintenance cost sharing activities are likely to also be high benefit activities.

The indirect costs of research material sharing are first and foremost linked to the quality assurance process outlined above. Flaws in publicly shared research materials can trigger more behavioral effects (shame, regret) but also reputation costs. Another important aspect are proprietary costs from scientific competition. While likely to be positive from the societal perspective, the indirect individual costs of being "scooped" for an interesting follow-up project can be sizable. Finally, there is also the possibility of unethical misuse of shared research materials. Malevolent researchers might use shared materials without proper attribution or use it with a strategy of "reverse p-hacking", meaning creating reproductions or replications that communicate a biased picture of the overall evidence to increase their chances for publication.

Rational researchers can be expected to voluntarily share research materials when their private benefits outweigh their private costs. Whether this happens will critically depend on whether the signaling mechanism outlined above unfolds. In a credence good market like research, signaling is a key mechanism to avoid market failure. Based on the discussion above, a critical aspect could be the proprietary cost of competition as one could argue that this cost is higher for high quality researchers with innovative ideas than for low quality researchers that will likely face limited demand for their materials. Also the risks of reputation costs or misuse of materials could be higher for high quality researchers. On the other hand, it can be argued that direct costs should be lower for high quality researchers because of superior skills. Finally, outside options should be generally higher for high quality researchers meaning that they face higher opportunity costs of their time but also less competitive pressure to keep their materials private.

The discussion should sound somewhat familiar to those that study voluntary disclosures by firms. What is different from our traditional firm-setting is that most researchers are not acting as entrepreneurs. Instead, they are employed by academic institutions or act as civil servants. This can to some extend mute the relevance of the benefits and the costs outlined above, at least for those researchers that hold tenured positions. In addition, the employing institutions might also interact with the decision whether or not to make research materials available. They are designing performance and tenure evaluation criteria that incentivize researchers. Also, they compensate researchers for their time and they provide them with infrastructure that influences and at times drastically reduces the direct costs of research material sharing. They might even hold property rights on the intellectual capital that their employees generate. Based on the public announcements of many leading academic institutions, it seems as if these are generally advocating the positive externalities of research material sharing. However, the incentive systems designed by these institutions appear at least in parts inconsistent with these announcements, as they are valuing individual research achievements in form of publications and only rarely publicly available research materials.

Finally, publishers and journals are influential for research material sharing decisions as well. Leading journals have established demanding data and code sharing guidelines, essentially transforming the voluntary disclosure decision to a mandatory disclosure requirement. While journals also compete for the best papers and it is an interesting question how this competition affects the dissemination of data and code sharing guidelines, it seems as if imposing mandatory disclosure rules is a classic response to overcome the diagnosed public good problem.

### Research Questions and Survey Design

Following the discussion above, we designed our survey to address three overlapping research questions.

1.  How does the accounting research community assess the reproducibility of their influential work?

2.  What are the views of the accounting research community about research material sharing and pre-registration, two prominent open science methods to increase the reproducibility of research?

3.  What are the (perceived) costs and benefits of research material sharing?

The first two questions are descriptive in nature. Assuming that the signaling mechanism above or other incentives have not succeeded in removing all insecurity about the reproducibility of influential accounting work, we expect to find some variation in our response with regard to research question one. More importantly, we want to compare the response of our sample with responses by researchers from other social science fields. The second question explores whether accounting researchers believe that research material sharing and pre-registration are viable strategies to increase the reproducibility of research. Again, comparing the views of accounting academics with the views of neighboring fields will be instrumental here.

The third question strives to understand the potential causes behind these assessments. Using a small experimental component, our survey tests whether two aspects of a hypothetical sharing request, its rationale (collaboration or control) and its scope (public versus private), affect the stated hypothetical likelihood of sharing. Our motivation for the treatments is that the rationale should shift the focus from benefits (collaboration) to costs (control) while the scope treatment should manipulate the salience of indirect costs. Besides studying the effect of these two experimental treatments, we also use verbal and structured survey response data to uncover perceived benefits and costs from research material sharing.

To verify whether the assessments provided by our survey participants are consistent with their behavior in the field, we also conduct an observational study that uses data from 100 authors of our target population to contrast the survey response with the actual research material sharing and pre-registration behavior in the field.

To ensure that our findings with respect to research questions one and two are comparable with related prior work, we adapted key questions from the Survey of Open Science Practices and Attitudes in the Social Sciences [@FLCPSWMBP_2023], published in Nature Communications. Similar to our survey, this study solicits the views of published researchers and PhD students on reproducibility, research material sharing, and pre-registration. However, there are key differences between the depth and target audience of our study and the study of @FLCPSWMBP_2023, warranting some adjustments. The latter lasted approximately 45 minutes and had a compensation element rewarding respondents for their participation. Our study, in contrast, is anonymous and has been designed to be completed in less than 10 minutes. It includes only mild incentivization in the form of a donation to UNICEF. Regardless of our adaptions, the ordering, phrasing, and answering scales of questions in our survey are similar. The adaptions enable us to considerably shorten the survey and to ensure that accounting researchers can relate the questions to their commonly accepted research practices. We pre-tested our adapted questions among a group of doctoral students specializing in accounting, which indicated no issues with the adapted questions. Appendix C includes a matching table providing an overview of every question and answer scale that we adapted from @FLCPSWMBP_2023.

Our survey has been designed using oTree [@CSW_2016]. The first stage of the survey includes questions about respondents' views on publicly sharing research materials, which we describe as comprising data, code, and research procedures. Specifically, we ask respondents about their familiarity with the public sharing of research materials and their views of its importance. We also ask them about their perceptions of the prevalence of sharing in the accounting research community, how often they tend to share, and whether they plan to share in the future. Conditional on whether respondents indicate that they do not always share, we also ask them for the reasons behind not sharing. Lastly, we ask respondents to estimate what percentage of accounting researchers favor sharing and their approximation of the percentage of editors of the top six accounting journals that believe sharing increases the value of a manuscript.

The second stage of the survey features a selection of questions about respondents' views on reproducibility and pre-registration, which we also borrowed and adapted from @FLCPSWMBP_2023. First, we ask respondents about their confidence in the reproducibility of accounting research. We also asked them about their subfield within the domain of accounting (i.e., Auditing, Financial Accounting and Reporting, Managerial Accounting, Taxation, Accounting information systems, and Other) and whether they believe the findings in their field are reproducible. Next, we ask the same questions for their methodological area (i.e., Analytical, Archival/Field, Experimental, Qualitative, Survey, and Other).

After eliciting respondents' perceptions on reproducibility, a set of questions explores their familiarity and experience with the practice of pre-registration. Lastly, we ask them about their perceptions of the trustworthiness of pre-registered research compared to non-pre-registered research. The survey ends with a few demographic questions. For a complete overview of the survey questions included in our study, please see Appendix A. All answers can be explored using our \href{https://jgassen.shinyapps.io/osacc/}{survey dashboard} and the data itself are available in our public \href{https://github.com/joachim-gassen/osacc/}{GitHub repository} .

### Survey Administration and Response Sample

To recruit respondents for our survey, we identified authors who published in the "Top 6" accounting journals, i.e., Accounting, Organizations and Society, Contemporary Accounting Research, Journal of Accounting and Economics, Journal of Accounting Research, Review of Accounting Studies, The Accounting Review, over the period 2016 to 2021. We chose these six accounting journals because they are commonly characterized as the leading accounting journals within the academic accounting community. We realize that this is not an exhaustive list and that there are divergent views about journals within the accounting community. Yet, we believe that, taken together, these journals provide a sufficiently broad target population of accounting academics.

We start compiling the list of authors from 2016 because this is the year that Open Science and, more specifically, the reproducibility of research became a more widespread topic of debate among academics within the social sciences and, presumably, accounting researchers [@OSC_2015; @B_2016; @CDF_2016; @M_2017]. After receiving approval from an ethical review board and guaranteeing the anonymity of the participating respondents, we sent out initial invitations and a total of three reminders. We allowed potential respondents to opt out of future emails every time we reached out to them. Please see Appendix A for the invitation emails and all reminders.

\centerline{--- Insert Table 1 about here ---}

Table 1 presents our sample selection procedures. Out of 2,660 potential respondents identified, 118 were excluded due to operational reasons.[^3] Additionally, 206 respondents' initial emails bounced, indicating unreachable email addresses. Among the remaining 2,336 successfully contacted accounting researchers, 391 initiated the study (15.4% of the total emails sent), 254 partially completed it (10.9% of the total emails sent), and 222 fully completed it (8.7% of the total emails sent). For our analyses, we employed a floating sample, which dynamically changes based on specific criteria, to maximize the use of the available data. For most descriptive analyses, we relied on the number of respondents who completed the entire study (n = 222). However, for certain descriptive analyses, depending on the specific question, we utilized a broader sample (n = 229) as seven respondents decided to leave out some of the demographic questions. Since our survey experiment component forms the first part of the study, we use the sample of respondents who completed the study up to that point (n = 254) for analysis.

\centerline{--- Insert Table 2 about here ---}

Table 2 displays response demographics and characteristics based on a sample of up to 229 observations. Panel A reveals that the majority of the respondents is male, 35 years or older (87.8%), specializing in Financial Accounting and Reporting or Managerial Accounting (70.3%), has Archival or Field Research as their primary research method (55.9%), and is from North America or Europe (91%). Panel B of Table 2 contrasts our respondent sample with our sampling population. Given the anonymity of our response and the data that we can collect about our population, we can only directly compare the geographic dispersion of our sample. Panel B reveals that, in particular, researchers from North America and Europe are over-represented, and researchers from Asia are under-represented in the sample. Given that response rate assessment is inherently challenging, we generally avoid speculating about potential reasons for this sampling imbalance. However, it seems likely that the study's all-European author team might have positively influenced the European response. While we control for geographical effects in our response (not-tabulated but available using our \href{https://jgassen.shinyapps.io/osacc/}{survey dashboard}) and find only mild variation across regions (researchers from North America tend to state higher sharing activities and preferences), we conclude from this sampling imbalance that our results mostly speak to the attitudes of accounting scholars from North America and Europe and that we can say very little about researchers located in Asia and other geographical regions that happen to have low response rates.

We further evaluate the representativeness of our sample by comparing it to a sample of published authors on the BYU accounting rankings platform [@SW_2023]. Specifically, the BYU accounting rankings rank individual authors based on classifications of peer-reviewed articles in 12 accounting journals since 1990. We contacted the team behind the BYU rankings, and they generously provided their data to us. The data includes `r format(nrow(byu), big.mark = ",")` authors and information on their methodologies as well as subfields. As these data are based on the papers contained in the BYU database as well as by self-classification by authors, one author can be associated with several subfields and methodologies in the BYU data. Therefore, we compare 2,114 classified methodologies in the BYU sample to 229 primary methodologies in our sample. Also, we compare 2,877 classified subfields in the BYU sample to 229 primary subfields in our sample.

\centerline{--- Insert Figures 1 and 2 about here ---}

Figure 1 reveals the frequency distributions of subfields in the BYU and our sample. We can observe some differences between these two distributions. First, our sample has disproportionately more mentions of authors publishing in financial accounting and reporting and managerial accounting than the BYU sample (49% versus 36%). As a result, we see disproportionately fewer authors working in the other subfields in our sample. A chi-square test confirms the frequencies of the two distributions significantly vary (chi-square = 36.415, two-tailed p-value \< 0.001). Figure 2 reveals the frequency distributions of the methodologies in the BYU and our sample. These two frequency distributions are very similar; a little over half of the observations fall into the Archival/Field category for both samples. About one-fifth of the sample falls into the experimental category, and the rest is comprised of analytical and other methodologies. A chi-square test does not reject the null that the two frequency distributions are the same (chi-square = 2.270, two-tailed p-value = 0.518).

## Findings

### Reproducibility

The first research question of our study is to explore how accounting researchers perceive the reproducibility of influential findings in their field. Table 3 provides a comprehensive overview of our respondents' answers to the question, "How confident are you that the influential findings in the area of accounting overall are reproducible?" The table also contrasts the answers with those of researchers in other disciplines as recorded by @FLCPSWMBP_2023, based on their question, "How confident are you that the influential research findings in \[discipline\] would replicate?" Notably, we adapted the wording in this question from "replicate" to "reproduce" to ensure clarity for our audience, recognizing the nuanced distinctions between these terms across various research methodologies. The term "reproducibility" is a focal point in the discussion across the social sciences, including accounting [@B_2016; @CDF_2016; @M_2017; @HLL_2000]. It implies the ability to recreate the results of prior studies using the same materials, procedures, and data as were used by the original researchers [@NSF_2015]. In contrast, the term "replication" has different interpretations for different fields. For instance, for experimental researchers, it refers to a distinct activity than reproduction (i.e., repeating an experimental study with new participants), while for archival researchers, it refers to a similar activity (i.e., re-analyzing the same or similar data using the same procedures). In turn, the definition that @FLCPSWMBP_2023 use for replication ("test of whether new studies on the same topic in a broadly similar sample support the same results.") seems relatively close to what we believe the common understanding of reproduction is in the area of accounting is. If anything, our notion of reproduction seems less demanding than the replication definition by @FLCPSWMBP_2023.

\centerline{--- Insert Table 3 about here ---}

The results in Table 3 reveal that accounting researchers generally hold a more critical view of the reproducibility of our influential findings. Compared to their peers in economics and sociology, accounting researchers exhibit a greater degree of doubt about the reproducibility of their influential work, with a difference of about half a point on a 5-point Likert scale. This difference is less stark when we compare accounting researchers to researchers in the fields of Political Science and Psychology. In a field fixed effect regression, in which we use Accounting as the baseline, we find that all four fixed field effects are statistically different from zero (two-tailed p-value \< 0.01). This skepticism is consistent with us accountants being more concerned about research quality hinting at the existence of a credence-good problem that open science methods are designed to address.

\centerline{--- Insert Figure 3 about here ---}

Figure 3 visually represents this skepticism, showing that 10.5% of accounting researchers are "Not at all confident" about the reproducibility of influential findings in their field. This percentage starkly contrasts the skepticism reported by economists (0.7%), political scientists (2.1%), psychologists (1.1%), and sociologists (1.3%), as found by @FLCPSWMBP_2023. In addition, this figure reveals that Accounting also has the lowest percentage of non-skeptics (i.e., researchers who are either moderately or extremely confident in the reproducibility of influential findings). That is, Accounting has about 39.3% non-skeptics, while Economics has 61.1%, Political Science has 45.8%, Psychology has 41.3%, and Sociology has 61.7%. The differences to Economics and Sociology, again, are the largest, while the differences to Political Science and Psychology are smaller.

To delve deeper, we examine the assessments of reproducibility across different methodologies within Accounting. Accounting is not only a diverse discipline in terms of orientation and topic but also in terms of the methodologies its researchers employ. We distinguish among five categories: Analytical, Archival/Field, Experimental, Qualitative, and Survey. We asked respondents which of those categories best described their main methodological area, enabling us to categorize them in the data. Table 4 presents their assessments regarding the reproducibility of influential findings across three panels. Panel A focuses on assessments of reproducibility within accounting as a whole, Panel B on assessments within their main methodological area, and Panel C on differences in assessments between accounting as a whole and their main methodological area.

\centerline{--- Insert Table 4 about here ---}

The findings in Table 4 Panel A suggest that analytical and qualitative researchers are more pessimistic and skeptical about the reproducibility of influential findings in accounting as a whole than accounting researchers using other methodologies. In particular, analytical researchers are significantly less confident about the reproducibility of influential findings in accounting in general compared to those using Archival/Field and Experimental categories (two-tailed p-value \< 0.100). Interestingly, when evaluating their own methodological areas, analytical researchers display the highest confidence levels, as their average confidence is statistically higher than those reported by the other four methodologies (two-tailed p-value \< 0.100), suggesting a discrepancy between their views on the field as a whole and their specific research approaches. Qualitative research reports significantly lower average confidence than Analytical, Archival/Field, and Experimental Research, albeit it is indistinguishable from the average confidence reported by those in the Survey category (two-tailed p-value \> 0.100).

Panel C presents the "assessment gap," which captures the difference in assessments of the reproducibility of influential findings in respondents' own methodological area versus accounting in general. A higher (lower) assessment gap indicates that respondents are more (less) confident in the reproducibility of influential findings in their methodological area compared to accounting in general. The findings of Panel C report that analytical researchers report the largest assessment gap, implying greater skepticism about the field's reproducibility as a whole compared to their own methodology. The average assessment gap reported for the Analytical category is also statistically different from the averages reported in the other four categories (two-tailed p-value \< 0.100). In contrast, the other methodological categories show no significant differences, suggesting a more uniform view across different research approaches within accounting.

### Sharing Research Materials - Familiarity and Importance

Accounting researchers so far show a disproportionate skepticism toward the reproducibility in their field relative to neighboring fields. Given this skepticism and following our discussion in section 2, one might expect accounting researchers to take a proactive stance on open science approaches, particularly regarding the sharing of research materials. Following the old proverb by Supreme Court Justice Luis Brandeis "sunlight is the best disinfectant" one would expect accountants to support increased transparency in research. The survey assesses this argument by asking respondents two questions about the practice of sharing research materials, which we define as the practice of sharing data, code, and research procedures. These two questions are also close adaptions of the survey done by @FLCPSWMBP_2023.

\centerline{--- Insert Table 5 about here ---}

The first question assesses respondents' familiarity with the practice of research materials sharing: "How familiar are you with the practice of publicly sharing research materials (i.e., data, code, and research procedures) for a completed study?"[^4] Table 5 Panel A shows that accounting researchers are generally less familiar with research materials sharing than their counterparts in other social sciences from the survey by @FLCPSWMBP_2023. Generally speaking, academics in the social sciences report being relatively aware of and familiar with the practice of research material sharing. However, the results do suggest that, compared to each of the other four fields in the social sciences, accounting researchers are significantly less familiar with the practice (two-tailed p-value \< 0.010). The second question captures respondents' perceived importance of research material sharing. [^5] Table 5 Panel B suggests a similar trend, whereby accounting researchers consider this practice less important than researchers in other social sciences (two-tailed p-value \< 0.010), a finding that underscores potential barriers to the adoption of open science principles.

\centerline{--- Insert Figure 4 about here ---}

Figure 4 graphically presents these findings, showing the frequencies along horizontal axes and centering them to a neutral midpoint. Only 39.6% of the accounting researchers deem research material sharing "very important" compared to higher percentages in other fields like Economics (51.7%), Political Science (48.1%), and Psychology (43.8%). Sociologists report a lower percentage of strong supporters (29.4%) than the accounting researchers in our sample, suggesting varied attitudes towards open science across the disciplines.

\centerline{--- Insert Table 6 about here ---}

We further examine attitudes, namely familiarity with and the perceived importance of research materials sharing across methodologies within accounting. Again, we distinguish between five methodological areas (i.e., Analytical, Archival/Field, Experimental, Qualitative, and Survey). Table 6 displays assessments of the practice of research material sharing using two panels. Similar to Table 5, Table 6 Panel A reports about the familiarity with the practice (i.e., "How familiar are you with the practice of publicly sharing research materials (i.e., data, code, and research procedures) for a completed study?") and Panel B reports about assessed importance ("To what extent do you believe that publicly sharing research materials (i.e., data, code, and research procedures) is important for the advancement of accounting research?").

Panel A shows that familiarity is larger for respondents using Archival and Field methods compared to respondents using Qualitative, Analytical, and Survey methods (two-tailed p-value \< 0.100). Moreover, we also find that respondents using Qualitative methods are less familiar with research material sharing compared to all four other categories (two-tailed p-value \< 0.100). Panel B shows that there are few significant differences in terms of the perceived importance of research material sharing across the methodologies. Only respondents in the Archival/Field category assess research material sharing as significantly more important than respondents using the Qualitative and Survey categories. Overall, these findings indicate that accounting researchers are, at the very least, not strong believers in the signaling value of research material sharing. This might be because they are simply relatively more unaware of the activity and/or they are assessing the benefits to be relatively lower or the costs to be relatively higher.

### Sharing Research Materials - Prescriptive versus Descriptive Views

Next, our survey focuses on the perceived prescriptive and descriptive importance of research material sharing. Table 7 Panel A reports data on the frequency of sharing, revealing that an impressive 57% of the respondents state that they at least sometimes publicly share their research materials. We also find that stated sharing tendencies are significantly lower for respondents in the Qualitative and Survey categories compared to each of the three other methodological areas, i.e., the respondents in the Analytical, Archival, and Experimental categories (two-tailed p-value \< 0.100).

\centerline{--- Insert Table 7 about here ---}

Table 7 Panel B reports answers to three questions measuring respondents' perceptions of how important others consider research material sharing to be (i.e., prescriptive and descriptive importance). Namely, these three questions ask (1) what percentage of accounting researchers in general favor research material sharing, (2) what percentage of editors of the top six accounting journals believe it increases the value of a manuscript, and (3) what percentage of accounting researchers publicly share research materials. The data in Panel B reveal that respondents believe about 44% of accounting researchers favor publicly sharing research materials, they believe 61% of the top-six accounting editors are in favor of publicly sharing research materials, and they believe 27% of accounting researchers actually share research materials.

\centerline{--- Insert Figure 5 about here ---}

To visualize the differences between these three questions, Figure 5 displays scatter box plots, one for each question and separately for each methodology. First, examining the three plots for the sample as a whole, the clear visual pattern emerges that the assessments of actual sharing behavior among accounting researchers lag behind assessments of whether accounting researchers favor sharing, which, in turn, lags behind assessment of whether editors value sharing. This pattern seems starker for respondents in the Archival/Field category. These findings suggest a contrast between the perceived (prescriptive) importance of research material sharing and the perceived (descriptive) actual sharing behavior. In other words: We believe that we should share and we also state that we do but we do not believe that others do.

### Pre-registration

Our survey has so far revealed that accounting researchers who participated in our survey are more skeptical about reproducibility than other social science areas. We also find that they are less familiar with research materials sharing as an open science practice that helps address reproducibility and that they find that practice less important. Besides research material sharing, our survey also covers a second Open Science practice: Pre-registration.

\centerline{--- Insert Table 8 about here ---}

Table 8 summarizes three questions about pre-registration. Panel A considers respondents' familiarity with the practice itself. We asked respondents: "How familiar are you with the practice of pre-registering hypotheses or analyses in advance of a study?"[^6] The results in Panel A suggest accounting researchers are equally familiar with the practice as researchers in Political Science and Psychology. They seem to be slightly more familiar than researchers in Economics and Sociology (two-tailed p-value \< 0.010). While this finding might appear surprising at first glance, one has to factor in that the practice of pre-registration is gaining traction over time, and our survey was fielded two years later than the second wave of our benchmark survey.

\centerline{--- Insert Figure 6 about here ---}

Panel B reports data on the first-time respondents pre-registered. It is based on the question: "Approximately when was the first time you pre-registered hypotheses or analyses in advance of a study?"[^7] The findings show a clear pattern: Accounting researchers started pre-registering their studies significantly later than researchers in other social science fields (two-tailed p-value \< 0.010). They started around 2018, compared to 2015-2016 for the other fields. This finding is also supported by Figure 6, which shows that the take-up of pre-registration in accounting over time lags behind the other social science areas.

Panel C examines trustworthiness, showing data on answers to the question: "How trustworthy do you find the work of scholars who tend to pre-register hypotheses or analyses as compared to the work of those who don't?"[^8] Results indicate no apparent differences in perceived trustworthiness in accounting compared to economics and political science (two-tailed p-value \> 0.100). However, the perceived increase in trustworthiness when pre-registering is less in accounting compared to psychology and sociology (two-tailed p-value \< 0.010). However, it should be noted that for this question, the comparability across fields is limited as we asked explicitly about trustworthiness, whereas @FLCPSWMBP_2023 referred more generally on a general opinion.

\centerline{--- Insert Table 9 about here ---}

To conclude our descriptive evidence on pre-registration, we examine differences in familiarity, usage, and trustworthiness of pre-registration across the different methodological areas within accounting. Table 9 displays respondents' frequencies using three panels. Similar to Table 8 Panel A, Panel A of Table 9 reports on familiarity, i.e., the question: "How familiar are you with the practice of pre-registering hypotheses or analyses in advance of a study?" Somewhat unsurprisingly, Panel A shows that qualitative accounting researchers are less familiar with the practice than accounting researchers belonging to the Archival, Experimental, and Survey categories (two-tailed p-value \< 0.100). Yet, there is no statistical difference between qualitative and analytical accounting researchers. Panel B reports on differences in usage, specifically, the question "`r vars$text[100]`", similar to the question featured in Table 8 Panel B. We find that accounting researchers in the experimental field report the highest usage of pre-registration, while none of the qualitative accounting researchers report using it. Lastly, Panel C reports differences in perceived trustworthiness. Specifically, it features answers to the question, "How trustworthy do you find the work of scholars who tend to pre-register hypotheses or analyses as compared to the work of those who don't?" which is identical to the question featured in Table 8 Panel C. We find no clear evidence of differences among the different methodologies in whether they trust pre-registered work more than work that has not been pre-registered.

### Usage of Open Science Methods - Observational Data

To contrast these perceptions with reality, we collect observational data on actual research material sharing behavior and the use of pre-registration behavior for a random country-stratified sample of `r nrow(authors)` authors from our target population. To assess whether these authors share research materials, we systematically search their homepages for evidence and also verify whether they have relevant online accounts linked to their names on platforms such as GitHub, Open Science Foundation, Dataverse, or Zenodo. If they do, we collect whether they share research materials over these channels. In addition, we use Scopus to identify all studies that these authors published since 2016 and verify whether these publications' DOI URLs reference research material sharing.[^9] We also assess various pre-registration platforms to assess whether our authors have used pre-registration. We only were able to identify for `r sum(by_author$has_prereged)` authors clear evidence that they have pre-registered a study. These cases are linked to the registered report initiative of JAR. Therefore, this reminder of this section solely focuses on research material sharing.

\centerline{--- Insert Table 10 about here ---}

Table 10 reports our findings based on this sample of `r nrow(authors)` authors and their `r nrow(papers)` papers. Panel A summarizes the data by author. We find that `r sum(by_author$has_made_resmat_available)` % of our authors share research materials. This number and its corresponding binomial confidence interval (30 to 49 %) are lower that the percentage of respondents that stated in the survey that they at least sometimes share (57 %) but a little bit higher than the prescriptive sharing frequency that our respondents attributed to other researchers (27 %).

\centerline{--- Insert Figure 7 about here ---}

As becomes apparent from Panel B, our authors only share research materials occasionally. Only 7.5 % of the paper DOI URLs reference research material sharing. As Figure 7 indicates, this share has mildly increased over time but even in recent years, the paper-level sharing percentage seems relatively modest.

The journal breakdown in Panel B provides clear evidence that the data and code sharing policies established by the Journal of Accounting Research and Management Science drastically influence research material sharing behavior. This also becomes apparent from Panel A. If one omits research material sharing via these journals, the percentage of sharing accounting researchers drops from 39 % to 23 %.[^10]

Overall, our observational data supports the survey findings. It seems that research material sharing only takes place sporadically in accounting and is often driven by journal requirements.

### Rationales Related to Research Material Sharing - Experimental Findings

We now turn to our third and final research question: An attempt to understand how our participants decide on whether to share research materials. As a first step, we use a small experimental design. For this, we create a hypothetical scenario. The scenario outlines that the participating researcher has just conducted a study and published it in an academic journal. Next, participants are informed that a fellow researcher has contacted them for access to their research materials. Next, we provide four pieces of information about the request made by a fellow researcher. In these statements, we manipulate two between-subjects factors: the rationale behind the request for research materials and the scope of sharing. First, we manipulate the rationale used by the fellow researcher requesting access to the researcher's materials. In the *collaboration rationale* treatment, the fellow researcher requests access to researcher materials to understand the work better and build on it. In the *control rationale* treatment, the request aims to verify and potentially correct the research. Second, we vary whether the fellow researcher requests the research materials to be shared publicly (i.e., *public scope*) or privately (i.e., *private scope*). Respondents were randomly assigned one stated rationale (*control* versus *collaboration*) and one sharing scope (*public* versus *private*). Their willingness to comply with the fellow researcher's request was then measured on a scale ranging from 0% (definitely not) to 100% (definitely). The hypothetical scenario was, like the survey, designed using oTree [@CSW_2016]. An overview of the manipulations and the materials can be found in Appendix B.

Building on our discussion in section 2, we predict that using a collaborative versus controlling rationale increases accounting researchers' willingness to share their research materials because it communicates a potential benefit rather than cost in terms of reputation (our first hypothesis). We also predict that a private sharing request would be more favorably received due to lower indirect costs triggered by lower reputation risks (our second hypothesis). Finally, we predict that the effect of a collaborative rationale is larger in the case of a private sharing request than a public sharing request (our third hypothesis). The conceptual reasoning behind this interaction effect is that, besides the additive effects of private sharing and a collaboration rationale, the perceived benefit of sharing under the collaboration rationale should be amplified when the scope of sharing is private rather than public. Namely, for a private scope, a collaboration rationale implies more direct benefits of sharing than for a public scope. We pre-registered our predictions and design at the \href{https://osf.io/4akyd}{Open Science Foundation}.

\centerline{--- Insert Table 11 about here ---}

Table 11 shows the findings of the hypothetical scenario. On average, we find a high willingness to share research materials in the hypothetical scenario and across different variations of rationale and scope (mean = 72%, s.d. = 28%). However, we do not find statistical support for our three predictions. First, Table 7 Column 1 reports no significant effect of using a *collaboration rationale* over a *control rationale* (two-tailed p-value \> 0.100). Moreover, Column 2 reveals no support for a main effect of *private scope* on respondents' willingness to share research materials (two-tailed p-value \> 0.100). Lastly, column 3 shows no support for an interaction effect between *collaboration rationale* and *private scope* (two-tailed p-value \> 0.100). In sum, although we find that respondents express a high willingness to share research materials in our hypothetical scenario, we do not find any evidence it is affected by the way the request for sharing is made. At least in our hypothetical scenario, it does not matter whether the request is made to make the materials private (rather than public) and whether the request is made to promote collaboration as a research community or to verify whether the research has been conducted appropriately.

### Rationales Related to Research Material Sharing - Survey Evidence

Also, given our experimental findings, we rely on two sources of survey data to shed light on our third research question. First, our survey contains a question asking respondents who report that they do not always publicly share research materials to explain which reasons make them withhold their research materials. Respondents are able to select multiple reasons.

\centerline{--- Insert Table 12 about here ---}

Table 12 Panel A displays the frequencies and percentages for each reason featured and provided in our survey. Aside from reporting the total frequencies for the entire sample, it also (again) splits them up into the five methodological fields (i.e., Analytical, Archival/Field, Experimental, Qualitative, and Survey). The most commonly stated reason that respondents gave for not sharing research materials is that the data are proprietary. 56.3% of all respondents give this reason, ranking first for all methodological areas. This is somewhat surprising because not all methodological areas regularly deal with proprietary data. In a shared second place are three more prominent reasons respondents provided, namely, (1) the time that it takes to make materials available (29.9% of all respondents), (2) the fear of being scooped when materials become publicly available (23.2% of all respondents), and (3) no need for sharing as their data are already publicly available (33.9% of all respondents). For these reasons, there are some notable differences across the different methodologies. For qualitative researchers, deidentification problems are a relatively important reason (45.8%). Analytical researchers, in contrast, consider the time that it takes to make materials available relatively less of a reason than researchers using the other methods (i.e., 9.5%).

Besides this multiple-choice question, we also ask our participants to reflect on their answers in our hypothetical sharing scenario. This second data source is interesting (a) because it provides unfiltered insights on the reasons for and against data sharing, and (b) it might be influenced by our experimental manipulations, despite that it did not affect the sharing likelihood in the hypothetical scenario. To gather these additional insights, we tasked a research assistant, who was not informed about the study's experimental design, to code the answers of the participants. The resulting categories, characterizing the answers and their respective frequencies, are presented in Panel B of Table 12. It also splits the frequencies by the different experimental cells of the experiment.

We first focus on the benefits of research material sharing. Panel B reveals that participants share research materials predominantly because they believe sharing enhances research quality. A typical example of such a statement is the following:

```{=tex}
\begin{displayquote}
"The purpose of research is to better understand the world, and replication (and scrutiny) of one's research is an essential element of developing a deep understanding of the world."
\end{displayquote}
```

While many participants refer to research in general when making this point, there are also responses that focus on the respondents' own work. These argue that research material sharing, by uncovering honest mistakes, can improve the quality of the sharing researcher's work. See, for example, the following statement:

```{=tex}
\begin{displayquote}
"Because if there is a shadow of a doubt, I pay the consequences. Moreover, I care about the results--so if they can be improved upon, I am all for it!."
\end{displayquote}
```

Another prominent argument provided by participants is that research material sharing helps other researchers. This argument is significantly more likely to be provided by participants who faced the collaboration treatment, providing evidence that the framing of the sharing situation affects the decision-making of our participants. We take this as evidence that our participants see the public and also private benefits of data and code sharing. The following statement illustrates the importance of research material sharing helping other researchers nicely:

```{=tex}
\begin{displayquote}
"I always make supporting information available to other researchers when requested. It may help them learn our process and expand upon our work."
\end{displayquote}
```

The main rationales against sharing are related to the materials being considered valuable intellectual property that can potentially be reused for other research projects. A typical example of such a statement reflecting the indirect proprietary costs of data sharing is:

```{=tex}
\begin{displayquote}
"There might have been substantial workload flown into collecting data that also will benefit follow-up projects. Giving away the data might harm future projects without remuneration for the collection effort."
\end{displayquote}
```
Another interesting aspect that does not become apparent from the pre-conditioned analysis in Panel A is that a sizable amount of participants indicate that they are concerned about the unethical use of their research materials. These concerns are related to other researchers using their work for "reverse p-hacking," trying to discredit prior findings to establish a contribution. An example of such a statement:

```{=tex}
\begin{displayquote}
"It sounds like they are trying to disprove me. They're on a witch hunt. There's a good chance they are doing this to a lot of people and are, themselves, going to suffer from a file drawer problem. If they re-run 100 people's studies and find that 1 doesn't replicate, that's the one they will publish and discredit and shame the researcher, ignoring the fact that it was likely due to chance, and not noting the 99 who got it right."
\end{displayquote}
```
But even without reverse p-hacking, research material sharing requests can create career concerns. An example:

```{=tex}
\begin{displayquote}
"I would want to share my materials but I would be nervous that I did something 'wrong'  and it would make me look bad if my study didn't replicate. I'm a junior faculty and wouldn't want to do anything to jeopardize my reputation or likelihood of getting tenure."
\end{displayquote}
```
Another aspect, which is also featured prominently in Panel A, is the time investment that is required to entertain research material sharing and potential follow-up requests. For example:

```{=tex}
\begin{displayquote}
"Many academics, including me, are constantly working under high pressure to deliver performance and short of time to fulfill a large range of commitments. It is simply impossible to entertain each and every such request, especially when replying to the first might lead to feeling the obligation to answer more questions from the requester in the future. A requester can be a relatively inexperienced graduate student exploring a doctoral topic, which has implications for how many questions s/he might have in the future. Thus, the propensity to reply to a request often depends on the status of the requester in the field and the anticipated consequences (or the lack of) of entertaining the request (or not)."
\end{displayquote}
```
The last aspect of this quote also hints at a rationale commonly featured in the "it depends" category: the requesting person's personality and reputation. An example of this is:

```{=tex}
\begin{displayquote}
"It depends on whether I know the person and their reputation in the research community. I think my reputation would be tarnished if I didn't provide but I would also be skeptical of providing someone I didn't know my hand-collected data and code that took years of work to produce."
\end{displayquote}
```
These statements indicate the potential for selective sharing among well-connected researchers, which would benefit an "in-crowd" of academics while at the same time unleveling the academic playing field and increasing inequality.

Taken together, we conclude from the rationale analysis that our participants give more consideration to potential sharing costs than to potential sharing benefits. While the sharing benefits become more prevalent with a collaboration framing, the sharing costs are prominent regardless of the hypothetical scenario. Also, while the multiple-choice data highlight direct cost arguments, the unstructured verbal response also features various indirect costs.

## Discussion and Conclusion

Informed by a voluntary diclosure themed discussion of the costs and benefits of open science activities in accounting, we conducted a survey among accounting researchers to explore the perceptions and attitudes towards open science in our field. Our results suggest that accounting researchers face an "open science dilemma." On the one hand, they are more skeptical about the reproducibility of their influential findings than other social science researchers are about their respective findings. On the other hand, they seem to be a little bit behind the curve in terms of research material sharing and research design pre-registration. If these conclusions are descriptive, what should our reaction be?

While we are not in a position to advise our profession on how to proceed, we feel that a two-pronged approach might be desirable. First, why are we so skeptical about our work? Is this an example of the anecdotal "professional skepticism" that leads us accountants to over-weight negative signals in general? Or do we indeed have good reasons to be concerned? The only way to find out is to take stock. @LSA_2024 suggest reproduction studies in accounting might not be rare, and a majority of those confirm prior results. However, we need more systematic evidence on whether our findings reproduce and whether they can be replicated using new data and settings. An impressive exercise in that regard is the recent large-scale reproduction attempt of studies published in Management Science [@FGHKO_2023]. This endeavor systematically reproduced almost 500 articles published in Management Science before and after the establishment of a Data and Code Policy in 2019. One interesting finding from that study is that accounting studies in Management Science are -if anything- more reproducible than studies in neighboring fields like finance. Based on this, it seems that professional skepticism and the resulting overly skeptical view on the reproducibility of our influential studies might indeed be partly responsible for our open science dilemma.

It would be insightful to compare the findings for Management Science, which also documents an impressive effect of their data and code policy, with data on studies published in the Journal of Accounting Research, again pre and post their data policy. These findings could, in turn, be compared with those for studies published in other leading journals that do not have a dedicated data and code sharing policy (yet). But not only systematic large-scale reproduction studies can update our priors about the reproducibility of our work. Replication games, potentially embodied in doctoral education, are an alternative means to assess the reproducibility of our work while at the same time helping researchers to grow their network and empirical skill sets alike. Taken together, we believe that the academic accounting profession should be well-positioned to take a leading role in verifying its output. In the end, assessing the reproducibility of scientific work is an assurance and auditing process, an area we know quite a bit about!

The second element of the two-pronged approach would be to promote research material sharing and pre-registration, potentially in combination with other open science measures that aim to increase the accessibility and reusability of scientific work. The economists among us might be quick to point out the incentive problems in this regard. Why should we do it? Does sharing code and data help to build academic careers? While some associative evidence from other fields indicates that papers that share their data are cited more often [@PDF_2007], currently, the answer to this is not clearly affirmative. Research material sharing requires skills and time, and, to the extreme, devoting this time might even be regarded as a negative signal by tenure and promotion committees, assuming that it crowds out time spent on producing "more tenurable output."

To us, this mindset seems to be the core of the problem. We believe that it is misguided as research material sharing constitutes a public good that is crucial for scientific progress. Think about a scientific computing world without open-source contributions. This would be a world without the GNU/Linux operating system and its tools, without Python, R, git, Latex, and the accompanying package ecosystems. In such a world, scientific advancements in areas like econometrics, machine learning, and natural language processing would be severely delayed if at all feasible. As our field builds on the insights of these areas, we all benefit from the open source/science paradigm in our everyday work.

Given that, one might ask whether research material sharing by accounting researchers can create similar positive externalities. As our competitive advantage stems more from institutional than methodological expertise, we, for example, are well-positioned to curate micro-level data based on firm disclosures. Prominent examples include word lists for sentiment analyses in financial settings [e.g., @LM_2011], topic model-based measures of specific firm risks [e.g., @H_2019], and industry classifications based on financial disclosures [e.g., @HP_2016].

A closely related area where we see the potential of accounting data to develop a substantial impact is the collection and pre-processing of firm-time level web and electronic filing data. For the former, web scraping of corporate webpages can provide informative data [e.g., @BBB_2021], and for the latter, in particular, the upcoming electronic filing requirements for sustainability reports hold promise. While researchers compete with established commercial data providers in this area, the need for research to study unstructured textual data makes it necessary and valuable to build research-focused data pipelines and products instead of simply trusting (and paying) commercial data vendors. Another aspect is that freely licensable research data can remove one of the most significant entry hurdles for empirical archival accounting researchers.

While these flagship projects are useful showcases that research material sharing in the accounting domain can also be impactful for other domains, they will likely not change the sharing landscape at large. To promote reproducible workflows overall, we need to make the case that adopting such a workflow not only contributes to the public good by making research materials accessible but also enhances research efficiency at an individual level. To be able to share research materials, such as code and data, researchers must adopt a workflow that integrates reusability aspects into project development. The data and code requirements of leading journals in economics, finance, and accounting are not only pushing the availability of research materials but also imposing substantial costs on researchers who have to provide more or less complete replication packages upon acceptance. As many researchers affected by these regulations will admit, transforming an empirical project from an internal "well it works" scripting exercise to a well-documented replication package that ideally works across multiple platforms is far from trivial. This process becomes significantly more manageable and enjoyable if an open science-oriented workflow is adopted throughout all stages of project development. Besides this, it also makes it easier for co-author teams to cooperate, and even the researcher's future self might be thankful.

As the lack of time is one of the biggest roadblocks to research material sharing in our survey, educational activities targeted at the PhD level teaching scientific computing and project management, in general, seem important. While well-funded research institutions might have the opportunity to outsource these skills to research professionals, the vast majority of researchers worldwide need to acquire them themselves. This is not necessarily undesirable, as having these skills enables one to develop more creative research projects than researchers unaware of what is feasible with state-of-the-art scientific computing. In the early days, many accounting PhD programs were implicit- or explicitly relying on scientific computing skills to be handed down from one PhD generation to another. It seems that these days are over and that dedicated scientific computing courses should become as central a component of scientific education in the quantitative social sciences as lab training is in the natural sciences.

Much of the above applies primarily to our discipline's empirical and quantitative areas. What about analytical and qualitative researchers and their studies? As it becomes apparent from our findings, their views on open science differ prominently from those of the "accounting mainstream." Analytical researchers seem to be much more confident in their work than others, potentially in line with theoretical work being easier to verify during the review process. Then again, to develop the impact that it deserves, analytical work not only needs to be reproducible but also accessible and reusable, ideally also by non-expert researchers from neighboring methods. Qualitative work, on the other hand, faces conditions that set it apart from typical quantitative work. First, its data is often collected under confidentiality, legally restricting future sharing. Second, its unique richness makes it inherently challenging to share across subjects, especially if re-identification issues are present. Third, related and maybe more epistemological in nature, the assumption of objectivity, which allows research materials to be detached from the individual researcher and reused by others, is questioned within the qualitative realm. This implies that qualitative scholars have to develop -and are developing [e.g., @M_2014; @FRV_2022]- their own understanding of what reproducibility means within their methodological paradigm.

Finally, fears and a general lack of trust in fellow researchers are also essential deterrents to adopting open science methods. This is human but unfortunate, also because it leads to subtle "home biases" when it comes to sharing, potentially increasing academic inequality and hindering scientific progress. While the fear of being "scooped" can be addressed by sharing materials later in the publication process or publicly claiming intellectual ownership of ideas via pre-registration and use of well-established repositories, fears with regards to future researchers pointing to deficiencies in one's work and/or a general lack of trust are more challenging to overcome. For both aspects, we need to work towards changing the error climate in the academic discipline. Additionally, it might be helpful to move away from the "policing" aspect of reproducibility and instead stress the collaborative public good character of open science activities. With them, we can help each other, promote our field, and make the research process even more fun!

\newpage

## References {.unnumbered}

\singlespacing

\small

::: {#refs}
:::

\doublespacing

\newpage

## Figures {.unnumbered}

### Figure 1: Sample Comparison between BYU and Our Sample - Subfields {.unnumbered}

```{r Figure1, fig.width=12, fig.height=9}
byu_fields_comp(byu, sd)
```

```{=tex}
\begin{singlespace}
\setlength{\parindent}{0em}
```
This figure compares the distribution of authors' subfields in the BYU sample to the distribution of respondents' self-reported primary subfield in our sample. In contrast to our survey, authors are classified in multiple subfields in the BYU sample. Therefore, this figure compares the 2,877 subfield classifications of the BYU sample to the 229 primary subfield classifications of our sample.

```{=tex}
\setlength{\parindent}{4em}
\end{singlespace}
\newpage
```
### Figure 2: Sample Comparison between BYU and Our Sample - Methods {.unnumbered}

```{r Figure2, fig.width=12, fig.height=9}
byu_methods_comp(byu, sd)
```

```{=tex}
\begin{singlespace}
\setlength{\parindent}{0em}
```
This figure compares the distribution of authors' methodologies in the BYU sample to the distribution of respondents' self-reported primary methodology in our sample. In contrast to our survey, authors are classified in multiple methodologies in the BYU sample. Therefore, this figure compares the 2,114 methodology classifications of the BYU sample to 229 primary methodology classifications of our sample.

```{=tex}
\setlength{\parindent}{4em}
\end{singlespace}
\newpage
```
### Figure 3: Reproducibility of Influential Findings Across the Social Sciences {.unnumbered}

```{r Figure3, fig.width=8, fig.height=5}
likert_plot(mr, "reprod", "demog_discipline", title_str = "How confident are you that the influential research findings in your area will replicate/reproduce?", llevels = vars$resp_label[80:84])
```

```{=tex}
\begin{singlespace}
\setlength{\parindent}{0em}
```
This figure is based on our survey data merged with the data collected by @FLCPSWMBP_2023. The question asked in our survey is "`r vars$text[80]`" while the original question used in the survey of @FLCPSWMBP_2023 is "How confident are you that the influential research findings in \[Discipline\] would replicate?". The data of @FLCPSWMBP_2023 has been collected in two waves (2018 and 2020) from a population of researchers located in the U.S. Our survey was conducted in 2022 among accounting researchers worldwide. Our survey and the merging process of the data are discussed in more detail in section 2 of the paper.

```{=tex}
\setlength{\parindent}{4em}
\end{singlespace}
\newpage
```
### Figure 4: Importance of Research Material Sharing Across the Social Sciences {.unnumbered}

```{r Figure4, fig.width=9, fig.height=5}
mr$sharing_important_rescaled = floor(mr$sharing_important)
likert_plot(mr, "sharing_important_rescaled", "demog_discipline", title_str = "To what extent do you believe that publicly sharing research materials is important?", llevels = vars$resp_label[49:53])
```

```{=tex}
\begin{singlespace}
\setlength{\parindent}{0em}
```
This figure is based on our survey data merged with the data collected by @FLCPSWMBP_2023. The question asked in our survey is "`r vars$text[49]`" while the original question used in the survey of @FLCPSWMBP_2023 is "To what extent do you believe that publicly posting (data or code \| research instruments) online is important for progress in \[Discipline\]?". The data of @FLCPSWMBP_2023 has been collected in two waves (2018 and 2020) from a population of researchers located in the U.S. Our survey was conducted in 2022 among accounting researchers worldwide. Our survey and the merging process of the data are discussed in more detail in section 2 of the paper.

```{=tex}
\end{singlespace}
\setlength{\parindent}{4em}
\newpage
```
### Figure 5: Sharing of Research Materials: Prescriptive versus Descriptive Views {.unnumbered}

```{r Figure5, fig.width=8, fig.height=6}
pres_desc_plot(sd)
```

```{=tex}
\begin{singlespace}
\setlength{\parindent}{0em}
```
This figure is based on our survey data and summarizes the findings from three separate questions. The first question, related to the preferences of editors is "`r vars$text[79]`". The second question, related to the preferences of other researchers, is "`r vars$text[78]`". The third question, related to the the actual behavior of other researchers, is "`r vars$text[54]`". Our survey is discussed in more detail in section 2 of the paper.

```{=tex}
\setlength{\parindent}{4em}
\end{singlespace}
```
\newpage

### Figure 6: Pre-Registration: Take-up over Time Across the Social Sciences {.unnumbered}

```{r Figure7, fig.width=8, fig.height=6}
pre_reg_over_time_plot(mr)
```

```{=tex}
\begin{singlespace}
\setlength{\parindent}{0em}
```
This figure is based on our survey data merged with the data collected by @FLCPSWMBP_2023. The question of our survey is "`r vars$text[102]`". The question used in the survey of @FLCPSWMBP_2023 is "Approximately when was the first time you pre-registered hypotheses or analyses in advance of a study?". The data of @FLCPSWMBP_2023 has been collected in two waves (2018 and 2020) from a population of researchers located in the U.S. Our survey was conducted in 2022 among accounting researchers worldwide. Our survey and the merging process of the data are discussed in more detail in section 2 of the paper.

```{=tex}
\setlength{\parindent}{4em}
\end{singlespace}
\newpage
```
```{=tex}
\renewcommand{\thetable}{\arabic{table}}
\setcounter{table}{0}
```
\newpage

### Figure 7: Share of Papers with Shared Research Materials over Time {.unnumbered}

```{r Figure6, fig.width=8, fig.height=6}
plot_resmat_sharing_over_cyear
```

```{=tex}
\begin{singlespace}
\setlength{\parindent}{0em}
```
This figure is based on `r nrow(papers)` papers, published since 2016 by a country-stratified random sample of `r nrow(authors)` authors from our target participant population. We collect the papers' DOIs from Scopus. For each paper, we observe whether research material availability is mentioned on the paper's DOI URL. The respective shares of papers are plotted by DOI creation year, along with their binomial confidence intervals. DOI creation years with less than 50 papers are excluded.

```{=tex}
\setlength{\parindent}{4em}
\end{singlespace}
\newpage
```
## Tables {.unnumbered}

### Table 1: Sample Composition {.unnumbered}

```{=tex}
\begin{singlespace}
\setlength{\parindent}{0em}
```
|                                 |                    N (%) |
|:--------------------------------|-------------------------:|
| Authors identified (Raw Sample) |                    2,660 |
| \- Excluded or missing          | 118 (4.4% of identified) |
| = Emails sent                   |                    2,542 |
| \- Emails bounced               |       206 (8.1% of sent) |
| = Emails received               |                    2,336 |
| Surveys Started                 |      391 (15.4% of sent) |
| Surveys Partially Completed     |      254 (10.9% of sent) |
| Surveys Fully Completed         |       222 (8.7% of sent) |

This table reports on our sample composition. Starting with a list of authors that we identified by Scopus to have published in the "Top 6" accounting journals (Accounting, Organizations and Society, Contemporary Accounting Research, Journal of Accounting and Economics, Journal of Accounting Research, Review of Accounting Studies, The Accounting Review) over the period 2016 to 2021, we excluded all authors with missing email information. We recorded any email bounce messages that we received but report the response rate based on the emails that we sent out. Our survey is discussed in more detail in section 2 of the paper.

```{=tex}
\setlength{\parindent}{4em}
\end{singlespace}
\newpage
```
### Table 2: Sample Demographics {.unnumbered}

```{=tex}
\begin{singlespace}
\setlength{\parindent}{0em}
```
#### Panel A: Sample Descriptives {.unnumbered}

```{r Tab2aDemographics, asis = TRUE}
extract_df <- function(dvar, title) {
	df <- create_demog_table(sd, dvar, df_only = TRUE)
	names(df) <- c("Characteristic", "n", "%")
	bind_rows(
		tibble(
			Characteristic = title,
			n = "",
			`%` = ""
		),
		df %>% mutate(Characteristic = paste0("- ", Characteristic))
	)
}

tab <- bind_rows(
	extract_df("demog_field", "Subfield"),
	extract_df("demog_method", "Method"),
	extract_df("demog_age", "Age Group"),
	extract_df("demog_gender", "Gender"),
	extract_df("demog_region", "Geographic Region")
)
gt(tab) %>% as_latex()

```

\newpage

```{=tex}
\setlength{\parindent}{4em}
\end{singlespace}
\newpage
```
```{=tex}
\begin{landscape}
\begin{singlespace}
\setlength{\parindent}{0em}
```
#### Panel B Regional Representativeness {.unnumbered}

```{r Tab2bGeoRep, asis = TRUE}
tab <- create_regions_rep_table(pop, sd)
tab_latex <- gt(tab) %>% 
	cols_align("right", -1) %>% 
	tab_spanner(label = "Population", columns = starts_with("Population")) %>%
	tab_spanner(label = "Sample", columns = starts_with("Sample")) %>%
	cols_label(
    ends_with(" n") ~ "n",
    ends_with(" %") ~ "%",
    ends_with("Rate") ~ "Regional RR"
	) %>%
		as_latex() 
# tab_latex[1] <- sub("\\\\toprule\\n", "", tab_latex)
tab_latex[1] <- sub("\\{Sample\\} &  ", "{Sample} &  \\\\multirowcell{2}{Regional \\\\\\\\ Response Rate}", tab_latex)
sub("& Regional RR", "& ", tab_latex)
```

This table details the demographics of our sample. As respondents were free not to answer any given question, the total number of answers vary by demographic. In Panel B, we contrast the geographical distribution of our sample with the geographical distribution of the survey population. For this, we use the affiliation of our authors as indicated by Scopus. For `r sum(is.na(pop$affil_country))` cases, the affiliation's country is unavailable on Scopus. Our survey is discussed in more detail in section 2 of the paper.

```{=tex}
\setlength{\parindent}{4em}
\end{singlespace}
\newpage
```
### Table 3: Reproducibility of Influential Findings Across the Social Sciences {.unnumbered}

```{=tex}
\begin{singlespace}
\setlength{\parindent}{0em}
```
How confident are you that the influential research findings in your field would reproduce/replicate?

```{r Tab3ReprodSocSciResp, asis = TRUE}
tab <- create_pct_cross_tab_table(
	mr, "reprod", "demog_discipline",
	resp_label =  get_resp_label("reprod_acc"), df_only = TRUE
)
gt(tab) %>% as_latex()
```

This table is based on our survey data merged with the data collected by @FLCPSWMBP_2023. The question asked in our survey is "`r vars$text[80]`" while the original question used in the survey of @FLCPSWMBP_2023 is "How confident are you that the influential research findings in \[Discipline\] would replicate?". The data of @FLCPSWMBP_2023 has been collected in two waves (2018 and 2020) from a population of researchers located in the U.S. Our survey was conducted in 2022 among accounting researchers worldwide. Our survey and the merging process of the data are discussed in more detail in section 2 of the paper. The asterisks report the results of an OLS regression regressing the response on a series of field fixed effects with Accounting acting as the base field. (\*/\*\*/\*\*\*) indicate two-sided significance at the (10%/5%/1%) significant level, respectively.

```{=tex}
\setlength{\parindent}{4em}
\end{singlespace}
\newpage
```
### Table 4: Reproducibility of Influential Findings: Assessment Across Methods {.unnumbered}

```{=tex}
\begin{singlespace}
\setlength{\parindent}{0em}
```
#### Panel A: General Assessment by Method {.unnumbered}

```{r Tab4aReprMethod, asis = TRUE}
tab <- create_pct_cross_tab_table(
	sd, "reprod_acc", "demog_method", include_total = TRUE,
	drop_levels = "Other", ctests = TRUE, df_only = TRUE
) 
gt(tab) %>% 
	cols_align("right", -1) %>% 
	as_latex()
```

#### Panel B: Method-Specific Assessment {.unnumbered}

```{r Tab4bReprMethod, asis = TRUE}
tab <- create_pct_cross_tab_table(
	sd, "reprod_method", "demog_method", include_total = TRUE,
	drop_levels = "Other", ctests = TRUE, df_only = TRUE
) 
gt(tab) %>% 
	cols_align("right", -1) %>% 
	as_latex()
```

\newpage

#### Panel C: Assessment Gap (Method - General) {.unnumbered}

```{r Tab4cReprMethod, asis = TRUE}
sd$reprod_method_delta <- as.numeric(sd$reprod_method) - as.numeric(sd$reprod_acc)
tab <- create_pct_cross_tab_table(
	sd, "reprod_method_delta", "demog_method", include_total = TRUE,
	drop_levels = "Other", 
	resp_label = sort(na.omit((unique(sd$reprod_method_delta)))),
	ctests = TRUE, add_num_labels = FALSE, df_only = TRUE 
) 
gt(tab) %>%
	cols_align("right", -1) %>% 
	as_latex()
```

This table is based on our survey data and two separate questions. The response presented in Panel A is based on the question "`r vars$text[80]`" while the response presented in Panel B is based on the question "`r vars$text[90]`". Panel C reports the individual level differences, measured in likert steps, between the answers given for Panel B and Panel A. In all method break-ups, but not for the totals, answers from three respondents who classified their method as "Other" are excluded. Methods where the answers differ at the two-sided 10%-level are indicated by their first two letters.

```{=tex}
\setlength{\parindent}{4em}
\end{singlespace}
\newpage
```
### Table 5: Sharing of Research Materials Across the Social Sciences {.unnumbered}

```{=tex}
\begin{singlespace}
\setlength{\parindent}{0em}
```
#### Panel A: Familiarity {.unnumbered}

```{r Tab5aSharingHeardSocSciResp, asis = TRUE}
tab <- create_pct_cross_tab_table(
	mr, "sharing_heard", "demog_discipline", 
	resp_label = c("No (0)", "Yes (1)"), 
	add_num_labels = FALSE, df_only = TRUE
) 
gt(tab) %>% 
	cols_align("right", -1) %>% 
	as_latex()
```

#### Panel B: Assessed Importance {.unnumbered}

```{r Tab5bSharingImportanceSocSciResp, asis = TRUE}
tab <- create_pct_cross_tab_table(
	mr, "sharing_important_rescaled", "demog_discipline", 
	resp_label = get_resp_label("sharing_important"), df_only = TRUE
) 
gt(tab ) %>% 
	cols_align("right", -1) %>% 
	as_latex()
```

```{=tex}
\setlength{\parindent}{4em}
\end{singlespace}
\newpage
```
### Table 6: Sharing of Research Materials Across Methods {.unnumbered}

```{=tex}
\begin{singlespace}
\setlength{\parindent}{0em}
```
#### Panel A: Familiarity {.unnumbered}

```{r Tab6aSharingFamiliarity, asis = TRUE}
tab <- create_pct_cross_tab_table(
	sd, "sharing_familiar", "demog_method", include_total = TRUE,
	drop_levels = "Other", ctests = TRUE, df_only = TRUE
) 
gt(tab) %>% 
	cols_align("right", -1) %>% 
	as_latex()
```

#### Panel B: Assessed Importance {.unnumbered}

```{r Tab6bSharingImportance, asis = TRUE}

tab <- create_pct_cross_tab_table(
	sd, "sharing_important", "demog_method", include_total = TRUE,
	drop_levels = "Other", ctests = TRUE, df_only = TRUE
) 
gt(tab) %>% 
	cols_align("right", -1) %>% 
	as_latex()
```

This table is based on our survey data and two separate questions. The response presented in Panel A is based on the question "`r vars$text[44]`" while the response presented in Panel B is based on the question "`r vars$text[49]`". In all method break-ups, but not for the totals, answers from three respondents who classified their method as "Other", and one observation from a person who chose not self-classify into a main method are excluded. Methods where the answers differ at the two-sided 10%-level are indicated by their first two letters.

```{=tex}
\setlength{\parindent}{4em}
\end{singlespace}
\end{landscape}
\newpage
```
### Table 7: Sharing of Research Materials: Prescriptive versus Descriptive Views {.unnumbered}

```{=tex}
\begin{singlespace}
\setlength{\parindent}{0em}
```
#### Panel A: Stated sharing frequency across methods {.unnumbered}

```{r Tab8aSharingUse, asis = TRUE}

tab <- create_pct_cross_tab_table(
	sd, "sharing_use", "demog_method", include_total = TRUE, 
	drop_levels = "Other", ctests = TRUE, df_only = TRUE
) 
gt(tab) %>% 
	cols_align("right", -1) %>% 
	as_latex()
```

#### Panel B: Sharing preference and sharing behavior believes across fields {.unnumbered}

```{r Tab8bSharingPreDes, asis = TRUE}
base_df <- sd %>%
	select(
		demog_method, sharing_perc_other_prefer, 
		sharing_perc_editor_prefer, sharing_perc_other_share
	) %>%
	na.omit()

stats_total <- base_df %>%
	pivot_longer(-1, names_to = "var", values_to = "values") %>%
	group_by(var) %>%
	summarise(
		Mean = sprintf("%.1f%%", mean(values)),
		`Std. Dev.` = sprintf("%.1f%%", sd(values)),
		`25%` = sprintf("%.1f%%", quantile(values, 0.25)),
		Median = sprintf("%.1f%%", median(values)),
		`75%` = sprintf("%.1f%%", quantile(values, 0.75)),
	) %>% 
	pivot_longer(-1, names_to = "stat", values_to = "Total") %>% 
	select(Total)

stats_fields <- base_df %>%
	filter(as.character(demog_method) != "Other") %>%
	group_by(demog_method) %>%
	pivot_longer(-1, names_to = "var", values_to = "values") %>%
	group_by(var, demog_method) %>%
	summarise(
		Mean = sprintf("%.1f%%", mean(values)),
		`Std. Dev.` = sprintf("%.1f%%", sd(values)),
		`25%` = sprintf("%.1f%%", quantile(values, 0.25)),
		Median = sprintf("%.1f%%", median(values)),
		`75%` = sprintf("%.1f%%", quantile(values, 0.75)),
	) %>% 
	pivot_longer(-1:-2, names_to = "stat", values_to = "values") %>%
	pivot_wider(
		id_cols = c(var, stat), 
		names_from = demog_method, values_from = "values"
	) %>% 
	ungroup() %>%
	select(-var) %>%
	rename(" " = stat)

stats <- bind_cols(stats_total, stats_fields) %>%
	select(" ", Total, everything())

tab <- rbind(
	c("General Preference for Sharing", rep("", 6)),
	stats[6:10,],
	c("Editor Preference for Sharing", rep("", 6)),
	stats[1:5,],
	c("Believes About Actual Sharing Behavior", rep("", 6)),
	stats[11:15,]
)

tab_latex <- gt(tab) %>% 
	cols_align("right", -1) %>% 
	as_latex()

tab_latex <- sub("General Preference for Sharing &  &  &  &  &  &", "\\\\multicolumn{7}{l}{General Preference for Sharing}", tab_latex)
tab_latex <- sub("Editor Preference for Sharing &  &  &  &  &  &", "\\\\multicolumn{7}{l}{Editor Preference for Sharing}", tab_latex)
sub("Believes About Actual Sharing Behavior &  &  &  &  &  &", "\\\\multicolumn{7}{l}{Believes About Actual Sharing Behavior}", tab_latex)

```

This table is based on our survey data and several separate questions. The response presented in Panel A is based on the question "`r vars$text[55]`" while the response presented in Panel B is based on the questions "`r vars$text[78]`" (general preference for sharing), "`r vars$text[79]`" (editor preference for sharing) and "`r vars$text[54]`" (actual sharing behavior). In all method break-ups, but not for the totals, answers from three respondents who classified their method as "Other", and one observation from a person who chose not to self-classify into a main method are excluded. Methods where the answers differ at the two-sided 10%-level are indicated by their first two letters. Please see Appendix C for an overview of how we adapted these three questions and their answer scales from @FLCPSWMBP_2023.

```{=tex}
\setlength{\parindent}{4em}
\end{singlespace}
\newpage
```
```{=tex}
\begin{singlespace}
\setlength{\parindent}{0em}
```
### Table 8: Pre-Registration Across the Social Sciences {.unnumbered}

#### Panel A: Awareness {.unnumbered}

```{r Tab11aPreRegAwareSocSciResp, asis = TRUE}
tab <- create_pct_cross_tab_table(
	mr, "prereg_heard", "demog_discipline", 
	resp_label = c("No (0)", "Yes (1)"), 
	add_num_labels = FALSE, df_only = TRUE
) 
gt(tab) %>% 
	cols_align("right", -1) %>% 
	as_latex()
```

#### Panel B: Take-up over Time {.unnumbered}

```{r Tab11bPreRegTimeSocSciResp, asis = TRUE}
tab <- create_pct_cross_tab_table(
	mr, "prereg_use_first_time", "demog_discipline", 
	resp_label = na.omit(sort(unique(mr$prereg_use_first_time))), 
	add_num_labels = FALSE, df_only = TRUE
)

# NOTE: This hack Needs to be checked manually whenever something 
# the underlying data changes - which it should not

up_to_2010 <- c(
	"<= 2010", 
	sum(as.numeric(substr(tab$Accounting[1:13], 1, 1))),
	sum(as.numeric(substr(tab$Economics[1:13], 1, 1))),
	sum(as.numeric(substr(tab$`Political Science`[1:13], 1, 1))),
	sum(as.numeric(substr(tab$Psychology[1:13], 1, 1))),
	sum(as.numeric(substr(tab$Sociology[1:13], 1, 1)))
)
up_to_2010[2:6] <- sprintf(
	"%s (%.1f %%)", 
	up_to_2010[2:6],
	100*as.numeric(up_to_2010[2:6])/as.numeric(tab[26,2:6])
)
tab <- rbind(up_to_2010, tab[14:28,])

gt(tab) %>% 
	cols_align("right", -1) %>% 
	as_latex()
```

```{=tex}
\newpage
\begin{landscape}
```
#### Panel C: Trustworthiness {.unnumbered}

```{r Tab11cPreRegTimeSocSciResp, asis = TRUE}
tab <- create_pct_cross_tab_table(
	mr, "prereg_opinion", "demog_discipline", 
	resp_label = get_resp_label("prereg_trust"), df_only = TRUE
) 
gt(tab ) %>% 
	cols_align("right", -1) %>% 
	as_latex()
```

This table is based on our survey data merged with the data collected by @FLCPSWMBP_2023. For Panel A, the question of our survey is "`r vars$text[95]`" where we code the response "Not at all familiar" as "No" and all other responses as "Yes." The question used in the survey of @FLCPSWMBP_2023 is "Have you ever heard of the practice of pre-registering hypotheses or analyses in advance of a study?". For Panel B, the question of our survey is "`r vars$text[102]`". The question used in the survey of @FLCPSWMBP_2023 is "Approximately when was the first time you pre-registered hypotheses or analyses in advance of a study?". For Panel C, the question of our survey is "`r vars$text[103]`". The question used in the survey of @FLCPSWMBP_2023 is "What is your opinion of pre-registering hypotheses or analyses?". The data of @FLCPSWMBP_2023 has been collected in two waves (2018 and 2020) from a population of researchers located in the U.S. Our survey was conducted in 2022 among accounting researchers worldwide. Our survey and the merging process of the data are discussed in more detail in section 2 of the paper. The asterisks report the results of an OLS regression regressing the response on a series of field fixed effects with Accounting acting as the base field. (\*/\*\*/\*\*\*) indicate two-sided significance at the (10%/5%/1%) significant level, respecively.

```{=tex}
\newpage
\setlength{\parindent}{0em}
```
### Table 9: Pre-Registration Across Methods {.unnumbered}

#### Panel A: Familiarity {.unnumbered}

```{r Tab12aPreRegFamiliarity, asis = TRUE}
tab <- create_pct_cross_tab_table(
	sd, "prereg_familiar", "demog_method", include_total = TRUE,
	drop_levels = "Other", ctests = TRUE, df_only = TRUE
) 
gt(tab) %>% 
	cols_align("right", -1) %>% 
	as_latex()
```

#### Panel B: Usage {.unnumbered}

```{r Tab12bPreRegUse, asis = TRUE}

tab <- create_pct_cross_tab_table(
	sd, "prereg_use", "demog_method", include_total = TRUE,
	drop_levels = "Other", 
	resp_label = c("No (0)", "Yes (1)"), 
	add_num_labels = FALSE, ctests = TRUE, df_only = TRUE
) 
# Adjust levels to be zero based.
tab[4,2:7] <- t(sprintf("%.2f", as.numeric(tab[4,2:7]) - 1))
tab[5,2:7] <- t(sprintf("%.2f", as.numeric(tab[5,2:7]) - 1))
gt(tab) %>% 
	cols_align("right", -1) %>% 
	as_latex()
```

\newpage

#### Panel C: Trustworthiness {.unnumbered}

```{r Tab12cPreRegTrust, asis = TRUE}

tab <- create_pct_cross_tab_table(
	sd, "prereg_trust", "demog_method", include_total = TRUE,
	drop_levels = "Other", ctests = TRUE, df_only = TRUE
) 
gt(tab) %>% 
	cols_align("right", -1) %>% 
	as_latex()
```

This table is based on our survey data and three separate questions. The response presented in Panel A is based on the question "`r vars$text[95]`", the response presented in Panel B is based on the question "`r vars$text[100]`", and the response presented in Panel C is based on the question "`r vars$text[103]`". In all method break-ups, but not for the totals, answers from three respondents who classified their method as "Other" are excluded. Methods where the answers differ at the two-sided 10%-level are indicated by their first two letters.

```{=tex}
\setlength{\parindent}{4em}
\end{landscape}
\end{singlespace}
\newpage
\begin{singlespace}
```
### Table 10: Observational Data on Research Material Sharing {.unnumbered}

#### Panel A: By Author {.unnumbered}

```{r Tab9aObservationalDataByAuthor, asis = TRUE}
gt(tab_by_author) %>% 
	cols_align("right", -1) %>% 
	as_latex()
```

#### Panel B: By Journal {.unnumbered}

```{r Tab9bObservationalDataByJournal, asis = TRUE}
gt(tab_by_journal) %>% 
	cols_align("right", -1) %>% 
	as_latex()

```

This table is based on on a country-stratified random sample of `r nrow(authors)` authors from our target participant population and `r nrow(papers)` papers that they published since 2016. We collect these publications' DOIs using Scopus. To identify research material sharing, we search homepages and the established open science platforms GitHub, Open Science Foundation (OSF), Dataverse, and Zenodo. For each paper, we observe whether research material availability is mentioned on the paper's DOI URL. The data collection manual for this process is available in this project's GitHub repository. In Panel A, the overall share of authors that we identify to share research materials is reported, along with its binomial confidence interval in brackets. We also report the share of authors that we document to provide research materials when we ignore the Journal of Accounting Research and Management Science, as both journals have mandatory data and code sharing policies in place. In addition, we report the share of authors that have accounts linked to their names on the open science platforms mentioned above. Panel B reports the paper-level data in total and separately for all journals for which we evaluated at least 20 papers.

```{=tex}
\setlength{\parindent}{4em}
\end{singlespace}
\newpage
```
### Table 11: Scenario Analysis: Hypothetical Sharing Likelihood {.unnumbered}

```{=tex}
\begin{singlespace}
\setlength{\parindent}{0em}
```
#### Panel A: Descriptive Statistics {.unnumbered}

```{r Tab7aExpDescriptives, asis = TRUE}
stats <- sd %>%
	group_by(exp_tment_control, exp_tment_private) %>%
	select(exp_tment_control, exp_tment_private,exp_pct_share) %>%
	summarise(
		N = sprintf("%d", n()),
		Mean = sprintf("%.1f%%", mean(exp_pct_share)),
		`S.D.` = sprintf("%.1f%%", sd(exp_pct_share)),
		Min = sprintf("%.1f%%", min(exp_pct_share)),
		`25%` = sprintf("%.1f%%", quantile(exp_pct_share, 0.25)),
		Median = sprintf("%.1f%%", median(exp_pct_share)),
		`75%` = sprintf("%.1f%%", quantile(exp_pct_share, 0.75)),
		Max = sprintf("%.1f%%", max(exp_pct_share))
	) %>%
	ungroup() %>%
	mutate(
		Rationale = c(rep("Collaboration", 2), rep("Control", 2)),
		Scope = c(rep(c("Public", "Private"), 2))
	) %>%
	select(-exp_tment_control, -exp_tment_private) %>%
	select(Rationale, Scope, everything())
tab <- create_pct_cross_tab_table(
	sd, "reprod_acc", "demog_method", 
	ctests = TRUE, df_only = TRUE
) 
gt(stats) %>% 
	cols_align("right", -1:-2) %>% 
	as_latex()
```

#### Panel B: Regression Results {.unnumbered}

```{r Tab7bExpRegResults, asis = TRUE}
sd$exp_tment_collab <- 1 - sd$exp_tment_control
modh1 <- lm(exp_pct_share ~ exp_tment_collab, data = sd)
modh2 <- lm(exp_pct_share ~ exp_tment_private, data = sd)
modh3 <- lm(exp_pct_share ~ exp_tment_collab*exp_tment_private, data = sd)
tab_latex  <- modelsummary(
	list(H1 = modh1, H2 = modh2, H3 = modh3), output = "gt",
	coef_rename = function(x) {
		x <- str_replace(x, fixed("exp_tment_collab"), "Collaboration rationale")
		x <- str_replace(x, fixed("exp_tment_private"), "Private scope")
		str_replace(x, fixed(":"), "×")
	},
	gof_map = list(
		list(raw = "adj.r.squared", clean = "Adjusted R2", fmt = 3),
		list(
			raw = "nobs", clean = "Number of observations", 
			fmt = function(x) format(x, big.mark=",")
		)
	),
	stars = c('*' = .1, '**' = 0.05, '***' = .01),
	estimate  = "{estimate}{stars}",
) %>% as_latex()

# The below is needed since some tex environment chock on the longtable environment
# See https://github.com/vincentarelbundock/modelsummary/issues/50
tab_latex[1] <- sub("\\(Intercept\\)", "Intercept", tab_latex)

sub("Adjusted R2", "\\\\midrule\\\\addlinespace[2.5pt]\nAdjusted R²", tab_latex)
```

This table reports the results from the pre-registered experimental design contained in our survey. Is is based on a hypothetical code and data sharing scenario were we manipulated the sharing rationale (collaboration versus control) and the sharing scope (private versus public). All experimental materials are detailed in the Online Appendix of this study. The outcome variable is the stated likelihood that the respondent would share their code and data under the hypothetical scenario. Panel A reports the descriptive statistics for this outcome, conditional for the different treatment cells. Panel B reports the results from OLS regressions of the outcome on the binary treatments (H1 and H2) and their interaction (H3).

```{=tex}
\setlength{\parindent}{4em}
\end{singlespace}
\newpage
```
### Table 12: Sharing of Research Materials: Rationales {.unnumbered}

```{=tex}
\begin{singlespace}
\setlength{\parindent}{0em}
```
#### Panel A: Reasons against Sharing (Close-Ended Question) {.unnumbered}

```{r Tab10aSharingReasons, asis = TRUE}
tab <- sd %>%
	select(starts_with("sharing_reason"), -starts_with("sharing_reason_other")) %>% 
	na.omit() %>%
	pivot_longer(everything(), names_to = "reason", values_to = "val") %>%
	group_by(reason) %>%
	summarise(
		sum = sum(val),
		Total = sprintf("%d (%.1f %%)", sum(val), 100*sum(val)/n())
	) %>%
	arrange(-sum) %>% 
	select(-sum) %>%
	left_join(
		sd %>%
			select(demog_method, starts_with("sharing_reason"), -starts_with("sharing_reason_other")) %>% 
			filter(demog_method != "Other") %>%
			na.omit() %>%
			pivot_longer(-1, names_to = "reason", values_to = "val") %>%
			group_by(demog_method, reason) %>%
			filter(sum(val) != 0) %>%
			summarise(
				stat = sprintf("%d (%.1f %%)", sum(val), 100*sum(val)/n())
			) %>%
			pivot_wider(id_cols = reason, names_from = demog_method, values_from = stat), 
		by="reason"
	) %>%
	mutate(
		reason =str_remove(reason, fixed("sharing_reason_")) %>% 
				str_replace_all(fixed("_"), " ") %>% str_to_sentence()
		) %>%
	rename(Reason = reason)

gt(tab) %>% 
	cols_align("right", -1) %>% 
	sub_missing(missing_text = "") %>%
	as_latex()
```

```{=tex}
\newpage
\begin{landscape}
```
#### Panel B: Rationales about Sharing (Open-Ended based on Hypothetical Scenario) {.unnumbered}

```{r Tab10bSharingRationales, asis = TRUE}
df <- sd %>% left_join(
	readRDS("../data/generated/verbal_response_coded.rds") %>% 
		select(-exp_reflect),
	by = "id"
) %>% mutate(
	verbal_gave_reasons = !is.na(verbal_n_reasons),
	verbal_mostly_positive = 1*(verbal_positive_reasons > verbal_negative_reasons),
	exp_tment_colab = 1 - exp_tment_control,
	exp_tment_group = 2*exp_tment_colab + exp_tment_private,
	demog_active_res = as.numeric(demog_active_res) - 1,
	demog_trust = (demog_trusting + demog_trustworthy)/2
)

reasons_overall <- df %>%
	filter(verbal_gave_reasons) %>%
	summarise(across(
		starts_with(c("verbal_pos_", "verbal_neg_", "verbal_dep_")), sum
	)) %>%
	pivot_longer(
		starts_with(c("verbal_pos_", "verbal_neg_", "verbal_dep_")), values_to = "n", names_to = "reason"
	) %>% mutate(
		reason = str_remove(reason, fixed("verbal_")),
		pct_all = n/sum(df$verbal_gave_reasons) 		
	)

reasons_by_tgroup <- df %>%
	filter(verbal_gave_reasons) %>%
	group_by(exp_tment_group) %>%
	summarise(across(
		starts_with(c("verbal_pos_", "verbal_neg_", "verbal_dep_")), 
		~ {sum(.x)/n()} 
	)) %>%
	pivot_longer(
		starts_with(c("verbal_pos_", "verbal_neg_", "verbal_dep_")), values_to = "pct", names_to = "reason"
	) %>% pivot_wider(
		names_from = "exp_tment_group", names_prefix = "pct_tg", values_from = "pct"
	) %>%
	mutate(
		reason = str_remove(reason, fixed("verbal_"))
	)

reasons <- reasons_overall %>% left_join(reasons_by_tgroup, by = "reason")

tab <- reasons %>%
	mutate(
		type = str_extract(reason, "pos|neg|dep"),
		reason = paste(
			"...",
			str_remove(reason, "pos_|neg_|dep_") %>% 
				str_replace_all(fixed("_"), " ") %>%
				tolower()
		),
		across(starts_with("pct"), ~ sprintf("%.1f %%", 100 * .x))
	) %>%
	mutate(other = (reason == "... other reasons")) %>%
	arrange(type, other, -n) %>%
	select(type, everything()) %>%
	select(-other)

names(tab) <- c(
	"Rationale", "Reason", "N", 
	"% Total", "% Control/Public", "% Control/Private", 
	"% Collab/Public", "% Collab/Private"
)

tab <- rbind(
	c("I share because it ... ", rep("", 6)),
	tab %>% filter(Rationale == "pos") %>% select(-Rationale),
	c("I do not share because ... ", rep("", 6)),
	tab %>% filter(Rationale == "neg") %>% select(-Rationale),
	c("It depends ... ", rep("", 6)),
	tab %>% filter(Rationale == "dep") %>% select(-Rationale)
)

tab_latex <- gt(tab) %>% 
	cols_align("right", -1) %>% 
	as_latex()

sub(
	"career concerns related to failed replications or corrections",
	"career concerns related to failed replications", tab_latex
)
```

This table is based on our survey data and two separate questions. The response presented in Panel A is based on the multiple choice answer to the close-ended question "`r vars$text[60]`". Participants answered this question if they expressed not always sharing their research materials. Panel B is based on the answer to the open-ended question that we pose as a follow-up in our experimental scenario: "`r vars$text[7]`". The free form answers to this question have been coded by an research assistant that had no information about the experimental design or the respective treatments. In Panel A, answers from two respondents who classified their method as "Other" are included in the Total column, but excluded from the methods break-up.

```{=tex}
\setlength{\parindent}{4em}
\end{landscape}
\end{singlespace}
\newpage
```
\setlength{\parindent}{0em}

## Online Supplement {.unnumbered}

The public GitHub repository at \href{https://github.com/joachim-gassen/osacc}{https://github.com/joachim-gassen/osacc} contains the data and the code of this project.

To explore our survey data in more depth, we encourage the use of our [online survey dashboard](https://jgassen.shinyapps.io/osacc/). Using this dashboard, you can explore the response to all survey questions, partitioning it by respondents' demographics.

\newpage

## Appendix A: Survey Administration and Instrument {.unnumbered}

```{=tex}
\begin{singlespace}
\setlength{\parindent}{0em}
\setlength{\parskip}{1em}
```
### Initial Invitation - November 21, 2022 {.unnumbered}

\[Title: Share your views on Accounting Research\]

Dear colleague,

We are conducting a study about sharing research materials, reproducibility, and pre-registration in accounting research. You have been chosen to provide your opinion on this subject because you are a published author in one of the leading accounting journals. We obtained your email address from publicly available sources.

Participating in our study takes around 7 minutes of your time. We will publicly share the anonymized data and results with the academic community in due course.

As a token of gratitude for your time, we will donate 1,000 € multiplied by the response rate to the United Nations International Children's Emergency Fund (UNICEF).

Click or copy and paste the URL below in your internet browser to participate in our study:\
\[URL\]

Should you have any other comments or questions, please feel free to contact us via \[EMAIL\].

We thank you for your consideration and participation.

Sincerely,

Joachim Gassen\
Humboldt-Universität zu Berlin\
TRR 266 Accounting for Transparency

Jacqueline Klug, Victor van Pelt\
WHU -- Otto Beisheim School of Management

Click here to unsubscribe from future e-mails. Click here to view our data protection guidelines.

Further information:

This study has been approved by the ethical review board of the WHU -- Otto Beisheim School of Management.

Participation in our study is entirely voluntary, and there are no consequences for you if you choose not to take part in it. We will ask for your consent to participate at the beginning of the study.

The data will only be analyzed and stored in anonymous form. This anonymous data will only be used for academic purposes and will be shared with other academics for research purposes.

\newpage

### Reminder 1 - November 28, 2022 {.unnumbered}

\[Title: Reminder: Share your views on Accounting Research\]

Dear colleague,

If you have not completed our **study** about **sharing research materials, reproducibility, and pre-registration in accounting research** yet, you still have time!

Participate in our **7-minute** study by following the URL below:

\[URL\]

As a token of gratitude for your time, we will **donate 1,000 € multiplied by the response rate** to the United Nations International Children's Emergency Fund (**UNICEF**).

Should you have any other comments or questions, please feel free to contact us via \[EMAIL\].

We thank you for your consideration and participation.

Sincerely,

Joachim Gassen\
Humboldt-Universität zu Berlin\
TRR 266 Accounting for Transparency

Jacqueline Klug, Victor van Pelt\
WHU -- Otto Beisheim School of Management

Click here to unsubscribe from future e-mails.

Click here to view our data protection guidelines.

Further information:

This study has been approved by the ethical review board of the WHU -- Otto Beisheim School of Management.

Participation in our study is entirely voluntary, and there are no consequences for you if you choose not to take part in it. We will ask for your consent to participate at the beginning of the study.

The data will only be analyzed and stored in anonymous form. This anonymous data will only be used for academic purposes and will be shared with other academics for research purposes.

\newpage

### Reminder 2 - December 5, 2022 {.unnumbered}

\[Title: Reminder: Share your views on Accounting Research\]

Dear colleague,

If you have not completed our **study** about **sharing research materials, reproducibility, and pre-registration in accounting research** yet, you still have time **until next Monday**. We will close the study after that date.

Participate in our **7-minute** study by following the URL below:

\[URL\]

As a token of gratitude for your time, we will **donate 1,000 € multiplied by the response rate** to the United Nations International Children's Emergency Fund (**UNICEF**).

Should you have any other comments or questions, please feel free to contact us via \[EMAIL\]

We thank you for your consideration and participation.

Sincerely,

Joachim Gassen\
Humboldt-Universität zu Berlin\
TRR 266 Accounting for Transparency

Jacqueline Klug, Victor van Pelt\
WHU -- Otto Beisheim School of Management

Click here to unsubscribe from future e-mails.

Click here to view our data protection guidelines.

Further information:

This study has been approved by the ethical review board of the WHU -- Otto Beisheim School of Management.

Participation in our study is entirely voluntary, and there are no consequences for you if you choose not to take part in it. We will ask for your consent to participate at the beginning of the study.

The data will only be analyzed and stored in anonymous form. This anonymous data will only be used for academic purposes and will be shared with other academics for research purposes.

\newpage

### Reminder 3 - December 12, 2022 {.unnumbered}

\[Title: Last Chance to Participate: Share your views on Accounting Research\]

Dear colleague,

If you have not completed our **study** about **sharing research materials, reproducibility, and pre-registration in accounting research** yet, this **Monday** is your **last chance to participate!** We will close the study by the end of today.

Participate in our **7-minute** study by following the URL below:

\[URL\]

As a token of gratitude for your time, we will **donate 1,000 € multiplied by the response rate** to the United Nations International Children's Emergency Fund (**UNICEF**).

Should you have any other comments or questions, please feel free to contact us via \[EMAIL\].

We thank you for your consideration and participation.

Sincerely,

Joachim Gassen\
Humboldt-Universität zu Berlin\
TRR 266 Accounting for Transparency

Jacqueline Klug, Victor van Pelt\
WHU -- Otto Beisheim School of Management

Click here to view our data protection guidelines.

Further information:

This study has been approved by the ethical review board of the WHU -- Otto Beisheim School of Management.

Participation in our study is entirely voluntary, and there are no consequences for you if you choose not to take part in it. We will ask for your consent to participate at the beginning of the study.

The data will only be analyzed and stored in anonymous form. This anonymous data will only be used for academic purposes and will be shared with other academics for research purposes.

\newpage

```{=tex}
\includepdf[scale=0.7, pages=1,pagecommand={\hypertarget{survey-instrument}{\subsection*{Survey Instrument}\label{survey-instrument}}\addcontentsline{toc}{subsubsection}{Survey Instrument}}]{materials/survey.pdf}
\includepdf[scale=0.7, pages=2-, pagecommand={}]{materials/survey.pdf}
```
## Appendix B: Scenario Manipulations and Instrument {.unnumbered}

### Manipulations {.unnumbered}

+---------------+-------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+--------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| Rationale     | Private Request                                                                                                                                                                                                         | Public Request                                                                                                                                                                                                                                   |
+---------------+-------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+--------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| Control       | You conducted a study and published it in an academic journal. Another researcher contacts you and asks for access to your research materials (i.e., the data, code, and research procedures). The researcher wants to: | You conducted a study and published it in an academic journal. Another researcher contacts you and asks you to publicly share your research materials (i.e., the data, code, and research procedures). The researcher wants to enable others to: |
|               |                                                                                                                                                                                                                         |                                                                                                                                                                                                                                                  |
|               | -   verify how your study was conducted,                                                                                                                                                                                | -   verify how your study was conducted,                                                                                                                                                                                                         |
|               |                                                                                                                                                                                                                         |                                                                                                                                                                                                                                                  |
|               | -   investigate whether your data and methodology are sound,                                                                                                                                                            | -   investigate whether your data and methodology are sound,                                                                                                                                                                                     |
|               |                                                                                                                                                                                                                         |                                                                                                                                                                                                                                                  |
|               | -   control whether your findings can be reproduced, and                                                                                                                                                                | -   control whether your findings can be reproduced, and                                                                                                                                                                                         |
|               |                                                                                                                                                                                                                         |                                                                                                                                                                                                                                                  |
|               | -   potentially conduct a new study that corrects your study.                                                                                                                                                           | -   potentially conduct a new study that corrects your study.                                                                                                                                                                                    |
+---------------+-------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+--------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| Collaboration | You conducted a study and published it in an academic journal. Another researcher contacts you and asks for access to your research materials (i.e., the data, code, and research procedures). The researcher wants to: | You conducted a study and published it in an academic journal. Another researcher contacts you and asks you to publicly share your research materials (i.e., the data, code, and research procedures). The researcher wants to enable others to: |
|               |                                                                                                                                                                                                                         |                                                                                                                                                                                                                                                  |
|               | -   understand how your study was conducted,                                                                                                                                                                            | -   understand how your study was conducted,                                                                                                                                                                                                     |
|               |                                                                                                                                                                                                                         |                                                                                                                                                                                                                                                  |
|               | -   learn from your data and methodology,                                                                                                                                                                               | -   learn from your data and methodology,                                                                                                                                                                                                        |
|               |                                                                                                                                                                                                                         |                                                                                                                                                                                                                                                  |
|               | -   contribute by generating new insights, and                                                                                                                                                                          | -   contribute by generating new insights, and                                                                                                                                                                                                   |
|               |                                                                                                                                                                                                                         |                                                                                                                                                                                                                                                  |
|               | -   potentially conduct a new study that builds on your study.                                                                                                                                                          | -   potentially conduct a new study that builds on your study.                                                                                                                                                                                   |
+---------------+-------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+--------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+

: {tbl-colwidths="\[20,40,40\]"}

```{=tex}
\end{singlespace}
```
```{=tex}
\includepdf[scale=0.7, pages=1,pagecommand={\hypertarget{scenario-instrument}{\subsection*{Instrument}\label{scenario-instrument}}\addcontentsline{toc}{subsubsection}{Scenario Instrument}}]{materials/scenario.pdf}
\includepdf[scale=0.7, pages=2-, pagecommand={}]{materials/scenario.pdf}
```
```{=tex}
\begin{landscape}
\begin{singlespace}
\begin{center}
```
## Appendix C: Comparison of our Survey with Social Science Survey (S3) {.unnumbered}

```{r AppCTab1, asis = TRUE}
df <- read_csv(
	"../data/external/s3_osacc_item_comparison.csv", 
	show_col_types = FALSE
)
gt(df[,1:7]) %>% 
	sub_missing() %>% 
	cols_width(
		`S3 Question` ~ pct(15),
		`Scale S3` ~ pct(15),
		`Our Question` ~ pct(15),
		`Our Scale` ~ pct(15),
		`Recoding` ~ pct(15)
	) %>%
	as_latex() %>%
	gt_add_header_mpage_table()
```

```{=tex}
\end{center}
\end{singlespace}
\end{landscape}
```

[^1]: The terms "reproduction" and "replication" are at times used as synonyms but have different meanings. Reproduction refers to achieving the same results of a prior study using the same materials (code, procedures, and data), while replication refers to duplicating the results of a of a prior study using the same code and procedures but with new data [@NSF_2015]. For a more in-depth discussion of the differences and related terminological debate, see @GFI_2016.

[^2]: We pre-registered this experimental part of our study. The pre-registration materials can be accessed [here](https://osf.io/frmzt/).

[^3]: We removed 41 authors because we found no email address, 41 authors because they appeared in the data twice with two author IDs but were the same person, 28 authors because they were a current member of one of our affiliated universities or academic research centers, 5 authors because they gave us advice for this study, 2 authors because they had passed away, and 1 author due to an incorrect email address.

[^4]: The corresponding question in @FLCPSWMBP_2023 is "Have you ever heard of the practice of publicly posting (data and code/study instruments) online for a completed study?" We deviated from this original question by using the term "research materials" and defining it as "sharing data, code, and research procedures" in our version of the question because we wanted to cover all aspects of this Open Science practice in one question.

[^5]: "To what extent do you believe that publicly sharing research materials (i.e., data, code, and research procedures) is important for the advancement of accounting research?" The corresponding question in @FLCPSWMBP_2023 is "To what extent do you believe that publicly posting (data or code \| research instruments) online is important for progress in \[Discipline\]?" We adapted this question for terminological consistency across our survey.

[^6]: The question used in the survey of @FLCPSWMBP_2023 is "Have you ever heard of the practice of pre-registering hypotheses or analyses in advance of a study?".

[^7]: The question used in @FLCPSWMBP_2023 was "Approximately when was the first time you pre-registered hypotheses or analyses in advance of a study?"

[^8]: The question used in the survey of @FLCPSWMBP_2023 is "What is your opinion of pre-registering hypotheses or analyses?". This question is quite broad while relying on a five-point Likert scale evaluating self-reported favorability (i.e., Not at all in favor, Moderately not in favor, Neither in favor nor against, Moderately in favor, Very much in favor). We chose to use a more specific question asking about whether pre-registration improves trustworthiness while relying on a five-point Likert scale (i.e., Much less trustworthy, Somewhat less trustworthy, Equally trustworthy, Somewhat more trustworthy, Much more trustworthy).

[^9]: The data collection manual for this process is available in this project's GitHub repository.

[^10]: Untabulated analyses based on small frequencies provide suggestive evidence that, overall, the sharing frequencies are higher for general economics journals than for journals in the areas of accounting and finance.