### 2025-03-13 docs/presentation.qmd
slide deck update

---
title: |
  | How Does the Accounting Research Community 
  | Think About Open Science?
subtitle: "" # we use this for the venue
author: Joachim Gassen, Jacqueline Klug, and Victor van Pelt
institute: "HU Berlin and WHU – Otto Beisheim School of Management"
date: today
date-format: "MMMM D, YYYY"
format: 
  beamer:
    latex_engine: xelatex # pdflatex creates rastered fonts
    slide_level: 3

classoption: "aspectratio=169"
bibliography: references.bib

header-includes:
- \usepackage{booktabs} 
- \usepackage{tabularx}
- \usepackage{multirow,makecell}
- \usepackage{array}
- \renewcommand\tabularxcolumn[1]{m{#1}}
- \usepackage{makecell}
- \usepackage{colortbl}
- \usepackage{adjustbox}
- \usepackage{tikz}
- \usepackage{siunitx}
- \usepackage{tabu}
- \usetikzlibrary{arrows,arrows.meta,calc,positioning,matrix}
- \input{materials/beamer_theme_trr266_16x9.sty}
---

```{r setup, include=FALSE, cache=F, message=F, warning=F, results="hide"}
knitr::opts_chunk$set(
  cache = FALSE, echo = FALSE, message = FALSE, warning = FALSE
)

suppressMessages({
  library(tidyverse)
  library(gt)
  library(kableExtra)
  library(modelsummary)
  library(showtext)
	library(tufte)
  devtools::source_url(
    "https://raw.githubusercontent.com/trr266/treat/main/code/R/theme_trr.R"
  )
})

set.seed(42)

source("../code/table_functions.R")
source("../code/figure_functions.R")

sd <- readRDS("../data/generated/survey_response.rds")
mr <- readRDS("../data/generated/merged_response.rds")
byu <- read_csv("../data/external/byu_author_data_anon.csv", show_col_types = FALSE)
vars <- read_csv("../data/external/variables.csv", show_col_types = FALSE)
pop <- read_csv(
	"../data/external/invited_authors_anon.csv", show_col_types = FALSE
)

authors <- read_csv("../data/external/obs_data_authors_anon.csv", show_col_types = FALSE)
papers <- read_csv("../data/external/obs_data_papers_anon.csv", show_col_types = FALSE)

source("../code/obs_data_analyses.R")

create_quote <- function(idval) {
	obs <- sd %>% filter(id == idval)
	tment_str <- paste0(
		ifelse(obs$exp_tment_control == 1, "Control", "Collab"), "/",
		ifelse(obs$exp_tment_private == 1, "Private", "Public")
	) 
	paste(obs$exp_reflect, "\\\\", quote_footer(paste("---", tment_str)))
}

# Needs \usepackage{adjustbox} in preamble
fit_gt_table_to_slide <- function(x, scale = 1) {
	if (scale == 1) repstr <- "\\\\resizebox{\\\\columnwidth}{!}{\\\\begin{tabular}"
	else repstr <- paste0("\\\\resizebox{", scale, "\\\\columnwidth}{!}{\\\\begin{tabular}")
  x <- sub("\\\\begin\\{longtable\\}", repstr, x)
  x <- sub("\\\\end\\{longtable\\}", "\\\\end{tabular}}", x)
  x
}

fit_gt_table_to_slide_long <- function(x, scale = 1) {
	if (scale == 1) repstr <- "\\\\resizebox*{!}{\\\\textheight}{\\\\begin{tabular}"
	else repstr <- paste0("\\\\resizebox*{!}{", scale, "\\\\textheight}{\\\\begin{tabular}")
  x <- sub("\\\\begin\\{longtable\\}", repstr, x)
  x <- sub("\\\\end\\{longtable\\}", "\\\\end{tabular}}", x)
  x
}

center_table <- function(x) {
	x[1] <- paste0("\\begin{center}", x[1], "\\end{center}")
	x
}

script_size <- function(x) {
	x[1] <- paste0("\\scriptsize", x[1], "\\normalsize")
	x
}

tiny_size <- function(x) {
	x[1] <- paste0("\\tiny", x[1], "\\normalsize")
	x
}
```

```{r InstallTRRFonts}
trr266_fonts_available <- all(c("Cambria", "Calibri Light") %in% font_families())

if (!trr266_fonts_available) {
  # On Mac with MS Office installation you need to add the Office
  # Fonts to the font path. The following depends on your 
  # installation but _should_ work in most cases

  office_font_path <- "/Applications/Microsoft Word.app/Contents/Resources/DFonts"
  if (Sys.info()["sysname"] == "Darwin") {
    if (dir.exists(office_font_path)) font_paths(office_font_path)
    else stop("MS Office font path not found")
  }

  rv <- tryCatch({
    font_add(family = "Cambria", regular = "Cambria.ttc")
    font_add(family = "Calibri Light", regular = "calibril.ttf")
  }, error = function(e) {
    message(sprintf("Failed to install TRR 266 fonts: %s", e))
    invisible(font_families())        
  })

  trr266_fonts_available <- all(c("Cambria", "Calibri Light") %in% rv)
  if (trr266_fonts_available) message("Successfully installed TRR 266 fonts") 
}
```

## Motivation: Estimating The Reproducibility of Psychological Science

\begin{center}
\includegraphics[height =0.7\textheight]{"materials/osc_2015.jpeg"} \\
Open Science Collaboration (2015, Science)
\end{center}

## Motivation: 1,500 Scientists Lift the Lid on Reproducibility

\begin{center}
\includegraphics[height =0.7\textheight]{"materials/baker_2016.jpeg"} \\
Baker (2016, Nature)
\end{center}

## Motivation: Survey of OS Practices and Attitudes in the Social Sciences

\begin{center}
\includegraphics[height =0.7\textheight]{"materials/ferguson_et_al_2023_uptake_over_time.png"} \\
Ferguson et al. (2023, Nature Communications)
\end{center}

## Our Study

Survey of authors that published in top six accounting journals over the period 2016-2021 to address three research questions:

1.  How does the accounting research community assess the reproducibility of their influential work?

2.  What are the views of the accounting research community about research material sharing and pre-registration, two prominent open science methods to increase the reproducibility of research?

3.  What are the (perceived) costs and benefits of research material sharing?

## Theoretical Framework

The acts of research material sharing and pre-registration can be understood as voluntary disclosure decisions.

- **Lack of unravelling**: Non-trivial direct costs and researchers might be unaware of their research' quality
- **Substantial indirect costs**: Competetive effects and reputational concerns
- **Limited private benefits**: Researchers are incentivized by publication success and less by research quality per se
- **Positive externalities**: Activities are likely to benefit other researchers without direct compensation for acting researcher (but: 'warm glow')

## Sample Composition {#sample}

|                                 |                    N (%) |
|:--------------------------------|-------------------------:|
| Authors identified (Raw Sample) |                    2,660 |
| \- Excluded or missing          | 118 (4.4% of identified) |
| = Emails sent                   |                    2,542 |
| \- Emails bounced               |       206 (8.1% of sent) |
| = Emails received               |                    2,336 |
| Surveys Started                 |      391 (15.4% of sent) |
| Surveys Partially Completed     |      254 (10.9% of sent) |
| Surveys Fully Completed         |       222 (8.7% of sent) |

\placebutton{skip}{email}{invitation email}

## Representativeness by Geographical Region

```{r Tab4GeoRep, asis = TRUE}
tab <- create_regions_rep_table(pop, sd)
tab_latex <- gt(tab) %>% 
	cols_align("right", -1) %>% 
	tab_spanner(label = "Population", columns = starts_with("Population")) %>%
	tab_spanner(label = "Sample", columns = starts_with("Sample")) %>%
	cols_label(
    ends_with(" n") ~ "n",
    ends_with(" %") ~ "%",
    ends_with("Rate") ~ "Regional RR"
	) %>%
	tab_options(column_labels.vlines.style = "hidden") %>%
	as_latex() %>%
	fit_gt_table_to_slide()
# tab_latex[1] <- sub("\\\\toprule\\n", "", tab_latex)
tab_latex[1] <- sub("\\{Sample\\} &  ", "{Sample} &  \\\\multirowcell{2}{Regional \\\\\\\\ Response Rate}", tab_latex)
sub("& Regional RR", "& ", tab_latex)
```

## Representativeness by Field

```{r FigureRespField, fig.width=12, fig.height=6, fig.align="center", fig.showtext = TRUE}
byu_fields_comp(byu, sd)
```

## Representativeness by Method

```{r FigureRespMethod, fig.width=9, fig.height=4.5, fig.align="center", fig.showtext = TRUE}
byu_methods_comp(byu, sd)
```

## Reproducibility Assessed Across the Social Sciences {#reproducibility}

```{r Figure1, fig.width=8, fig.height=3.5, fig.align="center", fig.showtext = TRUE}
p <- likert_plot(mr, "reprod", "demog_discipline", title_str = "How confident are you that the influential reserach findings in your area will replicate/reproduce?", llevels = vars$resp_label[80:84])
if (trr266_fonts_available) {
	p + theme(text = element_text(family = "Calibri Light", size = 7))
} else p + theme(
	title = element_text(size = 10),
	legend.text = element_text(size = 6)
)
```

\placebutton{skip}{terminology}{Terminology}

## Accounting Assessment by Method

```{r Tab4aReprMethod, asis = TRUE}
tab <- create_pct_cross_tab_table(
	sd, "reprod_acc", "demog_method", include_total = TRUE,
	drop_levels = "Other", ctests = TRUE, df_only = TRUE
)

gt(tab) %>% 
	cols_align("right", -1) %>% 
	as_latex() %>%
	fit_gt_table_to_slide()

```

## Assessment Gap (Method - General)

```{r Tab4cReprMethod, asis = TRUE}
sd$reprod_method_delta <- as.numeric(sd$reprod_method) - as.numeric(sd$reprod_acc)
tab <- create_pct_cross_tab_table(
	sd, "reprod_method_delta", "demog_method", include_total = TRUE,
	drop_levels = "Other", 
	resp_label = sort(na.omit((unique(sd$reprod_method_delta)))),
	ctests = TRUE, add_num_labels = FALSE, df_only = TRUE 
) 
gt(tab) %>%
	cols_align("right", -1) %>% 
	as_latex() %>%
	fit_gt_table_to_slide()
```

## Importance of Research Material Sharing Across the Social Sciences

```{r Figure2, fig.width=8, fig.height=3.5, fig.align="center", fig.showtext = TRUE}
p <- likert_plot(mr, "sharing_important", "demog_discipline", title_str = "To what extent do you believe that publicly sharing research materials is important?", llevels = vars$resp_label[49:53]) +
	theme(legend.text = element_text(size = 6))
if (trr266_fonts_available) {
	p + theme(
		text = element_text(family = "Calibri Light", size = 7),
		legend.text = element_text(size = 6)
	)
} else p + theme(legend.text = element_text(size = 6))

```

## Accounting Assessment of Sharing Importance by Method

```{r Tab6bSharingImportance, asis = TRUE}
tab <- create_pct_cross_tab_table(
	sd, "sharing_important", "demog_method", include_total = TRUE,
	drop_levels = "Other", ctests = TRUE, df_only = TRUE
) 
gt(tab) %>% 
	cols_align("right", -1) %>% 
	as_latex() %>%
	fit_gt_table_to_slide()
```

## Research Material Sharing: Prescriptive versus Descriptive Views

```{r Figure3, fig.width=8, fig.height=3.5, fig.align="center", fig.showtext = TRUE}
p <- pres_desc_plot(sd) + theme(legend.position = "right")
if (trr266_fonts_available) {
	p + theme(text = element_text(family = "Calibri Light", size = 7))
} else p
```

## Pre-Registration: Take-up over Time Across the Social Sciences

```{r Figure5, fig.width=8, fig.height=3.5, fig.align="center", fig.showtext = TRUE}
p <- pre_reg_over_time_plot(mr) + theme(legend.position = "right") 
if (trr266_fonts_available) {
	p + theme(text = element_text(family = "Calibri Light", size = 12))
} else p

```

## Pre-Registration: Opinion Across the Social Sciences

```{r Tab10cPreRegTimeSocSciResp, asis = TRUE}
tab <- create_pct_cross_tab_table(
	mr, "prereg_opinion", "demog_discipline", 
	resp_label = get_resp_label("prereg_trust"), df_only = TRUE
) 
gt(tab ) %>% 
	cols_align("right", -1) %>% 
	as_latex() %>%
	fit_gt_table_to_slide()
```

<!---
## Pre-Registration: Familiarity by Method

```{r Tab11aPreRegFamiliarity, asis = TRUE}
tab <- create_pct_cross_tab_table(
	sd, "prereg_familiar", "demog_method", include_total = TRUE,
	drop_levels = "Other", 
	ctests = TRUE, df_only = TRUE
) 
gt(tab) %>% 
	cols_align("right", -1) %>% 
	as_latex() %>%
	fit_gt_table_to_slide()
```

--->

## Pre-Registration: Usage by Method {#preregmethod}

```{r Tab11bPreRegUse, asis = TRUE}

tab <- create_pct_cross_tab_table(
	sd, "prereg_use", "demog_method", include_total = TRUE,
	drop_levels = "Other", resp_label = c("No (0)", "Yes (1)"), 
	add_num_labels = FALSE, ctests = TRUE, df_only = TRUE
) 
# Adjust levels to be zero based.
tab[4,2:7] <- t(sprintf("%.2f", as.numeric(tab[4,2:7]) - 1))
tab[5,2:7] <- t(sprintf("%.2f", as.numeric(tab[5,2:7]) - 1))
gt(tab) %>% 
	cols_align("right", -1) %>% 
	as_latex() %>%
	fit_gt_table_to_slide()
```

\placebutton{skip}{preregtrust}{trust in pre-registration by method}

## Observational Data: Author Level

```{r, asis = TRUE}
gt(tab_by_author) %>% 
	cols_align("right", -1) %>% 
	as_latex() %>%
	fit_gt_table_to_slide()
```

## Observational Data: Journal Level

```{r, asis = TRUE}
gt(tab_by_journal) %>% 
	cols_align("right", -1) %>% 
	as_latex() %>%
	fit_gt_table_to_slide()
```

## To Compare: Policy Effect for Management Science

\begin{center}
\includegraphics[width =0.7\textwidth]{"materials/fisar_et_al_2023_policy.jpeg"} \\
Fišar et al. (2023, Management Science)
\end{center}

## Observational Data: Paper-level Sharing over Time

```{r Figure4, fig.width=8, fig.height=3.5, fig.align="center", fig.showtext = TRUE}
p <- plot_resmat_sharing_over_cyear
if (trr266_fonts_available) {
	p + theme(text = element_text(family = "Calibri Light", size = 12))
} else p
```

## Sharing Determinants: Experimental Manipulation

You conducted a study and published it in an academic journal. Another researcher contacts you and asks you *(to publicly share | for access to)* your research materials (i.e., the data, code, and research procedures). The researcher wants *(others)* to:

\begin{columns}
\begin{column}{0.5\textwidth}
\begin{itemize}
\item verify how your study was conducted,
\item investigate whether your data and methodology are sound,
\item control whether your findings can be reproduced, and
\item potentially conduct a new study that corrects your study.                                \end{itemize}

\end{column}
\begin{column}{0.5\textwidth}
\begin{itemize}
\item understand how your study was conducted,
\item learn from your data and methodology,
\item contribute by generating new insights, and
\item potentially conduct a new study that builds on your study.                               
\end{itemize}
\end{column}
\end{columns}

## Costs and Benefits of Research Material Sharing: Experimental Findings

```{r, asis = TRUE}
sd$exp_tment_collab <- 1 - sd$exp_tment_control
modh1 <- lm(exp_pct_share ~ exp_tment_collab, data = sd)
modh2 <- lm(exp_pct_share ~ exp_tment_private, data = sd)
modh3 <- lm(exp_pct_share ~ exp_tment_collab*exp_tment_private, data = sd)
tab_latex  <- modelsummary(
	list(H1 = modh1, H2 = modh2, H3 = modh3), output = "gt",
	coef_rename = function(x) {
		x <- str_replace(x, fixed("exp_tment_collab"), "Collaboration rationale")
		x <- str_replace(x, fixed("exp_tment_private"), "Private scope")
		str_replace(x, fixed(":"), "×")
	},
	gof_map = list(
		list(raw = "adj.r.squared", clean = "Adjusted R2", fmt = 3),
		list(
			raw = "nobs", clean = "Number of observations", 
			fmt = function(x) format(x, big.mark=",")
		)
	),
	stars = c('*' = .1, '**' = 0.05, '***' = .01),
	estimate  = "{estimate}{stars}",
) %>% as_latex()

# The below is needed since some tex environment chock on the longtable environment
# See https://github.com/vincentarelbundock/modelsummary/issues/50
tab_latex[1] <- sub("\\(Intercept\\)", "Intercept", tab_latex)

sub("Adjusted R2", "\\\\midrule\\\\addlinespace[2.5pt]\nAdjusted R²", tab_latex)
```

## Reasons against Sharing (Close-Ended Question)

```{r Tab9aSharingReasons, asis = TRUE}
tab <- sd %>%
	select(starts_with("sharing_reason"), -starts_with("sharing_reason_other")) %>% 
	na.omit() %>%
	pivot_longer(everything(), names_to = "reason", values_to = "val") %>%
	group_by(reason) %>%
	summarise(
		sum = sum(val),
		Total = sprintf("%d (%.1f %%)", sum(val), 100*sum(val)/n())
	) %>%
	arrange(-sum) %>% 
	select(-sum) %>%
	left_join(
		sd %>%
			select(demog_method, starts_with("sharing_reason"), -starts_with("sharing_reason_other")) %>% 
			filter(demog_method != "Other") %>%
			na.omit() %>%
			pivot_longer(-1, names_to = "reason", values_to = "val") %>%
			group_by(demog_method, reason) %>%
			filter(sum(val) != 0) %>%
			summarise(
				stat = sprintf("%d (%.1f %%)", sum(val), 100*sum(val)/n())
			) %>%
			pivot_wider(id_cols = reason, names_from = demog_method, values_from = stat), 
		by="reason"
	) %>%
	mutate(
		reason =str_remove(reason, fixed("sharing_reason_")) %>% 
				str_replace_all(fixed("_"), " ") %>% str_to_sentence()
		) %>%
	rename(Reason = reason)

gt(tab) %>% 
	cols_align("right", -1) %>% 
	sub_missing(missing_text = "") %>%
	as_latex() %>%
	fit_gt_table_to_slide()
```

## Rationales about Sharing (Experimental Evidence)

```{r Tab9bSharingRationales, asis = TRUE}
df <- sd %>% left_join(
	readRDS("../data/generated/verbal_response_coded.rds") %>% 
		select(-exp_reflect),
	by = "id"
) %>% mutate(
	verbal_gave_reasons = !is.na(verbal_n_reasons),
	verbal_mostly_positive = 1*(verbal_positive_reasons > verbal_negative_reasons),
	exp_tment_colab = 1 - exp_tment_control,
	exp_tment_group = 2*exp_tment_colab + exp_tment_private,
	demog_active_res = as.numeric(demog_active_res) - 1,
	demog_trust = (demog_trusting + demog_trustworthy)/2
)

reasons_overall <- df %>%
	filter(verbal_gave_reasons) %>%
	summarise(across(
		starts_with(c("verbal_pos_", "verbal_neg_", "verbal_dep_")), sum
	)) %>%
	pivot_longer(
		starts_with(c("verbal_pos_", "verbal_neg_", "verbal_dep_")), values_to = "n", names_to = "reason"
	) %>% mutate(
		reason = str_remove(reason, fixed("verbal_")),
		pct_all = n/sum(df$verbal_gave_reasons) 		
	)

reasons_by_tgroup <- df %>%
	filter(verbal_gave_reasons) %>%
	group_by(exp_tment_group) %>%
	summarise(across(
		starts_with(c("verbal_pos_", "verbal_neg_", "verbal_dep_")), 
		~ {sum(.x)/n()} 
	)) %>%
	pivot_longer(
		starts_with(c("verbal_pos_", "verbal_neg_", "verbal_dep_")), values_to = "pct", names_to = "reason"
	) %>% pivot_wider(
		names_from = "exp_tment_group", names_prefix = "pct_tg", values_from = "pct"
	) %>%
	mutate(
		reason = str_remove(reason, fixed("verbal_"))
	)

reasons <- reasons_overall %>% left_join(reasons_by_tgroup, by = "reason")

tab <- reasons %>%
	mutate(
		type = str_extract(reason, "pos|neg|dep"),
		reason = paste(
			"...",
			str_remove(reason, "pos_|neg_|dep_") %>% 
				str_replace_all(fixed("_"), " ") %>%
				tolower()
		),
		across(starts_with("pct"), ~ sprintf("%.1f %%", 100 * .x))
	) %>%
	mutate(other = (reason == "... other reasons")) %>%
	arrange(type, other, -n) %>%
	select(type, everything()) %>%
	select(-other)

names(tab) <- c(
	"Rationale", "Reason", "N", 
	"% Total", "% Control/Public", "% Control/Private", 
	"% Collab/Public", "% Collab/Private"
)

tab <- rbind(
	c("I share because it ... ", rep("", 6)),
	tab %>% filter(Rationale == "pos") %>% select(-Rationale),
	c("I do not share because ... ", rep("", 6)),
	tab %>% filter(Rationale == "neg") %>% select(-Rationale),
	c("It depends ... ", rep("", 6)),
	tab %>% filter(Rationale == "dep") %>% select(-Rationale)
)

tab_latex <- gt(tab) %>% 
	cols_align("right", -1) %>% 
	as_latex()

sub(
	"career concerns related to failed replications or corrections",
	"career concerns related to failed replications", tab_latex
) %>%
	fit_gt_table_to_slide(0.9) %>% center_table()
```

## Quotes about Sharing Benefits: Helping Others
\begin{quote}
`r create_quote(295)`
\end{quote}
\vspace{1cm}
\begin{quote}
`r create_quote(403)`
\end{quote}

## Quotes about Sharing Benefits: Helping Yourself

\begin{quote}
`r create_quote(249)`
\end{quote}
\vspace{1cm}
\begin{quote}
`r create_quote(186)`
\end{quote}

## Quotes about Sharing Costs: Direct Costs

\begin{quote}
`r create_quote(154)`
\end{quote}

## Quotes about Sharing Costs: Valuable Intellectual Property

\begin{quote}
`r create_quote(47)`
\end{quote}

## Quotes about Sharing Costs: Unethical Usage Concerns

\begin{quote}
`r create_quote(169)`
\end{quote}

## Quotes about Sharing Benefits/Costs: Depends on Researcher

\begin{quote}
`r create_quote(317)`
\end{quote}

## Summary

We seem to face an **"accounting open science dilemma"**

- We are skeptical about the reproducibility of our work and at the same time
- somewhat behind the curve with regards to research material sharing and pre-registration.

Our findings suggest that weighting direct and indirect costs higher than private and public benefits might be a rationale for this behavior.

## Discussion

What now? We suggest a two-pronged approach

- Take stock of the reproducibility of our work
- Promote incentive-compatible open science approaches

## Taking stock: Distributions of Test Statistics in Archival Accounting Studies

\begin{center}
\includegraphics[height =0.7\textheight]{"materials/chang_gao_li_2025.jpeg"} \\
Chang, Gao, and Li (2025, JAR)
\end{center}

## Taking stock: Relatively speaking...

\begin{center}
\includegraphics[width =0.7\textwidth]{"materials/fisar_et_al_2023_fields.jpeg"} \\
Fišar et al. (2023, Management Science)
\end{center}

## Taking stock: Reproducibility Across Methods

\begin{center}
\includegraphics[width =0.8\textwidth]{"materials/fisar_et_al_2023_methods.jpeg"} \\
Fišar et al. (2023, Management Science)
\end{center}

## Promote Incentive-Compatible Open Science Approaches

- Use flagship projects to show that sharing can boost academic careers
- Demonstrate that scientific programming skills (i) enhance efficiency, (ii) open up avenues for creative research ideas, and (iii) make it easier to share.
- Change the error climate and move from the "policing" to a "collaboration" narrative for open science!

Interesting aspect for further study: Role of open science in theoretical/analytical and qualitative work (e.g., Moravcsik PolSci 2014; Freese et al. SocSciRes 2022)

## Reasons Against Sharing and Their Counter Arguments

\begin{center}
\includegraphics[height =0.7\textheight]{"materials/gomes_et_al_2022_sharing_concerns.jpeg"}\\
Gomes et al. (2022, Proceedings of the Royal Society Series B)
\end{center}

## Thank You!

\begin{center}
For additional descriptive analyses of our survey data, check out our \href{https://jgassen.shinyapps.io/osacc/}{survey dashboard}.

All code and data of this project are available on \href{https://github.com/joachim-gassen/osacc}{GitHub}.
\end{center}

# Backup

## Invitation Email {#email}

\begin{small}

Subject: Share your views on Accounting Research

Dear colleague,

We are conducting a study about sharing research materials, reproducibility, and pre-registration in accounting research. You have been chosen to provide your opinion on this subject because you are a published author in one of the leading accounting journals. We obtained your email address from publicly available sources.

Participating in our study takes around 7 minutes of your time. We will publicly share the anonymized data and results with the academic community in due course.

As a token of gratitude for your time, we will donate 1,000 € multiplied by the response rate to the United Nations International Children's Emergency Fund (UNICEF).

[...]

\end{small}

\placebutton{return}{sample}{Return}

## Some Terminology (Dreber and Johannesson, 2024) {#terminology}

Types of reproducibility:

- **Computational reproducibility**: Results of original study can be reproduced based on data and code posted or provided by the original authors.	
- **Recreate reproducibility**:	Results of original study can be reproduced based on the information in the paper and access to the same raw data source.	
- **Robustness reproducibility**:	Results og original study are robust to alternative reasonable analytical decisions on the same data.

Types of replicability:

- **Direct replicability**:	Results of original studies can be repeated on new data using the same research design and analysis as the original study.	
- **Conceptual replicability**:Results of original study can be repeated on new data using an alternative research design and/or analysis to test the same hypothesis.

\placebutton{return}{reproducibility}{Return}

## Pre-Registration: Trust Effect by Method {#preregtrust}

```{r Tab11cPreRegTrust, asis = TRUE}

tab <- create_pct_cross_tab_table(
	sd, "prereg_trust", "demog_method", include_total = TRUE,
	drop_levels = "Other", ctests = TRUE, df_only = TRUE
) 
gt(tab) %>% 
	cols_align("right", -1) %>% 
	as_latex() %>%
	fit_gt_table_to_slide()
```

\placebutton{return}{preregmethod}{Return}